---
title: "EI_thresholds"
output: html_document
---

Getting old packages from previous R versions
```{r}
# Get a list of packages from the old R version
old_lib <- "C:/Users/tamj/AppData/Local/R/win-library/4.4"  # adjust to your old R path
old_packages <- installed.packages(lib.loc = old_lib)[, "Package"]

# Reinstall them in the new R version
install.packages(old_packages)


install.packages("pak")
pak::pkg_install(old_packages)

```


# setup
```{r}
#system setup
require(tidyverse)
require(azmpdata)
require(marea)
require(gratia)
require(gradientForest)
require(mgcv)
```

# gather indicators and pressures
```{r}

marea::eco_indicators

#separate by region
# indicators used in Irvine et al. 2025 reasoning was that they didn't use any indicators calculated that had landings (not available at set level), MargalefRichness_ALL, Heips_ALL, LargeFishIndicator, MeanLengthAbundance, Biomass_Piscivore, CCondition_Finfish, CCondition_MBenthivore, CCondition_Zoopiscivore, CCondition_Planktivores, CCondition_LBenthivore, CVBiomass, MeanLifespan, BiomassTL2, Biomass_Skates, Biomass_All, Biomass_Flatfish, Biomass_Gadoids, Biomass_invertebrates

nafo_ei_4VN<-eco_indicators[eco_indicators@data$region=="4VN"]
nafo_ei_4VS<-eco_indicators[eco_indicators@data$region=="4VS"]
nafo_ei_4W<-eco_indicators[eco_indicators@data$region=="4W"]
nafo_ei_4X<-eco_indicators[eco_indicators@data$region=="4X"]
nafo_ind_4X<-nafo_ei_4X@data |> 
    select(margalefrichness=MargalefRichness_ALL_value,
           heips=Heips_ALL_value,
           lfi=LargeFishIndicator_value,
           mla=MeanLengthABUNDANCE_value,
           biomasspiscivore=BIOMASS_PISCIVORE_value
           )


ESS_ei<-eco_indicators[eco_indicators@data$region=="ESS"]
WSS_ei<-eco_indicators[eco_indicators@data$region=="WSS"]

# pull environmental indicators and fishing from azmpdata?
# nafo level, SST,Bottom Temp,bottom temperature anomalies, sst_anomalies,  Cal fin, zooplankton, Integrated chl, Nitrate (stations, lines?)
nafo_bt_4VN<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="4VN"]
nafo_bt_4VS<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="4VS"]
nafo_bt_4W<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="4W"]
nafo_bt_4X<-eco_indicators[eco_indicators@data$region=="4X"]
nafo_pres_4X<-nafo_ei_4X@data |> 
    select(landall=landings_ALL_value, 
           landforage=landings_FORAGE.L_value,
           fpall=FP_ALL_value,
           fpgadoids=FP_GADOIDS_value #no spaces or underscores!
           )

nafo_sst_4X<-azmpdata::Derived_Annual_Broadscale |>
        filter(area=="4XSS")|>
        select(YEAR=year,SST=sea_surface_temperature_from_satellite)

nafo_land_4X<-eco_indicators[eco_indicators@data$region=="4X"]
  
# Shelf level indicators, NAO_w, AMO, Bloom stats
  ESS_bt<-marea::azmp_bottom_temperature[azmp_bottom_temperature@data$region=="ESS"]
  WSS_bt<-marea::azmp_bottom_temperature[azmp_bottom_temperature@data$region=="WSS"]
  nao<-marea::nao
  amo<-marea::amo
  

```


# Step 1.  run GAM and GLMs for each indicator x pressure combo
```{r}
library(mgcv)
library(gratia)

# Initialize model list
model_list <- list()

# Loop to fit and store both GAM and GLM models
for (a_col in names(nafo_ind_4X)) {
  for (b_col in names(nafo_pres_4X)) {
    tryCatch({
      df <- data.frame(
        response = nafo_ind_4X[[a_col]],
        predictor = nafo_pres_4X[[b_col]]
      )
      df <- na.omit(df)
      
      if (nrow(df) < 10 || var(df$predictor) == 0 || var(df$response) == 0) next
      
      # Fit GAM
      gam_model <- gam(response ~ s(predictor), data = df, method = "REML",
                       family = tw())
      
      # Fit GLM (linear predictor)
      glm_model <- gam(response ~ predictor, data = df, method = "REML",
                      family = tw())
      
      # Save both models
      model_name <- paste(a_col, b_col, sep = "_vs_")
      model_list[[paste0(model_name, "_GAM")]] <- gam_model
      model_list[[paste0(model_name, "_GLM")]] <- glm_model
      
    }, error = function(e) {
      message("Error fitting models for ", a_col, " vs ", b_col, ": ", e$message)
    })
  }
}

```

# Step 2. Pull model results and create a data table of model results
```{r}
# Initialize comparison results data frame
comparison_results <- data.frame()

# Loop through model_list to extract diagnostics
for (model_name in names(model_list)) {
  tryCatch({
    model <- model_list[[model_name]]
    model_summary <- summary(model)
    
    # Determine model type and variable names
    parts <- strsplit(model_name, "_vs_")[[1]]
    response <- parts[1]
    predictor_and_type <- parts[2]
    
    # Separate predictor and model type
    predictor <- sub("_.*$", "", predictor_and_type)
    model_type <- sub("^.*_", "", predictor_and_type)
    
    # Extract diagnostics
    aic_val <- AIC(model)
    r_squared <- model_summary$r.sq
    edf <- if ("edf" %in% names(model_summary)) model_summary$edf[1] else NA
    gcv_score <- model$gcv.ubre
    
    # Extract p-value depending on model type
    p_value <- tryCatch({
      if (model_type == "GAM") {
        model_summary$s.table[1, "p-value"]
      } else if (model_type == "GLM") {
        model_summary$p.table["predictor", "Pr(>|t|)"]
      } else {
        NA
      }
    }, error = function(e) NA)
    
    # Store results
    comparison_results <- rbind(comparison_results, data.frame(
      Response = response,
      Predictor = predictor,
      Model_Type = model_type,
      AIC = aic_val,
      R_squared = r_squared,
      EDF = edf,
      GCV_UBRE = gcv_score,
      p_value = p_value,
      stringsAsFactors = FALSE
    ))
    
  }, error = function(e) {
    message("Error extracting diagnostics for ", model_name, ": ", e$message)
  })
}

# View comparison table
print(comparison_results)
```

# Model selection proceedure selecting lower GCV scores from the comparison_results table
```{r}
library(dplyr)

# Compare GAM vs GLM for each Response-Predictor pair
model_selection <- comparison_results %>%
  group_by(Response, Predictor) %>%
  filter(Model_Type %in% c("GAM", "GLM")) %>%
  arrange(GCV_UBRE) %>%
  slice(1)  # Select model with lowest GCV


print(model_selection)

```

# create plots 
```{r}
library(gratia)
library(ggplot2)
library(patchwork)

plot_list <- list()

for (i in seq_len(nrow(model_selection))) {
  row <- model_selection[i, ]
  response <- row$Response
  predictor <- row$Predictor
  model_type <- row$Model_Type
  
  full_model_name <- paste0(response, "_vs_", predictor, "_", model_type)
  model <- model_list[[full_model_name]]
  
  if (!is.null(model)) {
    if (model_type == "GAM") {
      # Use gratia::draw for GAMs
      p <- draw(model) +
        labs(
          title = paste("GAM:", response, "vs", predictor),
          x = predictor,
          y = response
        )
    } 
    
    # use ggplot2 for GLMs
  else if (model_type == "GLM") {
    df <- model$model
    df$fit <- fitted(model)
    df$se <- sqrt(model$scale) * predict(model, type = "link", se.fit = TRUE)$se.fit
  
  # Compute confidence bounds (±1 SD)
  df$upper <- df$fit + df$se
  df$lower <- df$fit - df$se
  
  p <- ggplot(df, aes_string(x = "predictor", y = "response")) +
    geom_point(alpha = 0.6) +
    geom_line(aes(y = fit), color = "blue", linewidth = 1) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "blue", alpha = 0.2) +
    labs(
      title = paste("GLM:", response, "vs", predictor),
      x = predictor,
      y = response
    ) +
    theme_minimal()
}

    
    plot_list[[full_model_name]] <- p
  }
}

# Combine and display plots
combined_plot <- wrap_plots(plot_list, ncol = 4)
print(combined_plot)

```




# Step 1.  run GAM and GLMs for each indicator x pressure combo
```{r}

library(mgcv)
library(gratia)

# Initialize model list
model_list <- list()

# Loop to fit and store both GAM and GLM models
for (a_col in names(nafo_ind_4X)) {
  for (b_col in names(nafo_pres_4X)) {
    tryCatch({
      df <- data.frame(
        response = nafo_ind_4X[[a_col]],
        predictor = nafo_pres_4X[[b_col]]
      )
      df <- na.omit(df)
      
      if (nrow(df) < 10 || var(df$predictor) == 0 || var(df$response) == 0) next
      
      # Fit GAM
      gam_model <- gam(response ~ s(predictor), data = df, method = "REML",
                       family = tw())
      
      # Fit GLM (linear predictor)
      glm_model <- gam(response ~ predictor, data = df, method = "REML",
                      family = tw())
      
      # Save both models
      model_name <- paste(a_col, b_col, sep = "_vs_")
      model_list[[paste0(model_name, "_GAM")]] <- gam_model
      model_list[[paste0(model_name, "_GLM")]] <- glm_model
      
    }, error = function(e) {
      message("Error fitting models for ", a_col, " vs ", b_col, ": ", e$message)
    })
  }
}

```

# Step 2. Pull model results and create a data table of model results
```{r}
# Initialize comparison results data frame
comparison_results <- data.frame()

# Loop through model_list to extract diagnostics
for (model_name in names(model_list)) {
  tryCatch({
    model <- model_list[[model_name]]
    model_summary <- summary(model)
    
    # Determine model type and variable names
    parts <- strsplit(model_name, "_vs_")[[1]]
    response <- parts[1]
    predictor_and_type <- parts[2]
    
    # Separate predictor and model type
    predictor <- sub("_.*$", "", predictor_and_type)
    model_type <- sub("^.*_", "", predictor_and_type)
    
    # Extract diagnostics
    aic_val <- AIC(model)
    r_squared <- model_summary$r.sq
    edf <- if ("edf" %in% names(model_summary)) model_summary$edf[1] else NA
    gcv_score <- model$gcv.ubre
    
    # Extract p-value depending on model type
    p_value <- tryCatch({
      if (model_type == "GAM") {
        model_summary$s.table[1, "p-value"]
      } else if (model_type == "GLM") {
        model_summary$p.table["predictor", "Pr(>|t|)"]
      } else {
        NA
      }
    }, error = function(e) NA)
    
    # Store results
    comparison_results <- rbind(comparison_results, data.frame(
      Response = response,
      Predictor = predictor,
      Model_Type = model_type,
      AIC = aic_val,
      R_squared = r_squared,
      EDF = edf,
      GCV_UBRE = gcv_score,
      p_value = p_value,
      stringsAsFactors = FALSE
    ))
    
  }, error = function(e) {
    message("Error extracting diagnostics for ", model_name, ": ", e$message)
  })
}

# View comparison table
print(comparison_results)
```

# Model selection proceedure selecting lower GCV scores from the comparison_results table
```{r}
library(dplyr)

# Compare GAM vs GLM for each Response-Predictor pair
model_selection <- comparison_results %>%
  group_by(Response, Predictor) %>%
  filter(Model_Type %in% c("GAM", "GLM")) %>%
  arrange(GCV_UBRE) %>%
  slice(1)  # Select model with lowest GCV


print(model_selection)

```

# create plots 
```{r}
library(gratia)
library(ggplot2)
library(patchwork)

plot_list <- list()

for (i in seq_len(nrow(model_selection))) {
  row <- model_selection[i, ]
  response <- row$Response
  predictor <- row$Predictor
  model_type <- row$Model_Type
  
  full_model_name <- paste0(response, "_vs_", predictor, "_", model_type)
  model <- model_list[[full_model_name]]
  
  if (!is.null(model)) {
    if (model_type == "GAM") {
      # Use gratia::draw for GAMs
      p <- draw(model) +
        labs(
          title = paste("GAM:", response, "vs", predictor),
          x = predictor,
          y = response
        )
    } 
    
    # use ggplot2 for GLMs
  else if (model_type == "GLM") {
    df <- model$model
    df$fit <- fitted(model)
    df$se <- sqrt(model$scale) * predict(model, type = "link", se.fit = TRUE)$se.fit
  
  # Compute confidence bounds (±1 SD)
  df$upper <- df$fit + df$se
  df$lower <- df$fit - df$se
  
  p <- ggplot(df, aes_string(x = "predictor", y = "response")) +
    geom_point(alpha = 0.6) +
    geom_line(aes(y = fit), color = "blue", linewidth = 1) +
    geom_ribbon(aes(ymin = lower, ymax = upper), fill = "blue", alpha = 0.2) +
    labs(
      title = paste("GLM:", response, "vs", predictor),
      x = predictor,
      y = response
    ) +
    theme_minimal()
}

    
    plot_list[[full_model_name]] <- p
  }
}

# Combine and display plots
combined_plot <- wrap_plots(plot_list, ncol = 4)
print(combined_plot)

```


```{r}
library(gratia)
library(ggplot2)

# Loop through GAM models in model_list
for (model_name in names(model_list)) {
  if (grepl("_GAM$", model_name)) {
    model <- model_list[[model_name]]
    
    # Compute derivatives of the smooth term
    deriv <- derivatives(model, term = "predictor", partial_match =TRUE)
    
    # Identify threshold points where derivative crosses zero
    zero_crossings <- deriv$predictor[which(diff(sign(deriv$derivative)) != 0)]
    
    # Plot derivative with threshold points
    p <- draw(deriv) +
      geom_vline(xintercept = zero_crossings, linetype = "dashed", color = "red") +
      labs(
        title = paste("Derivative Analysis:", model_name),
        x = "Predictor",
        y = "Derivative"
      )
    
    print(p)
  }
}

```

