---
title: "EI_thresholds"
output: html_document
---



# setup

```{r}
#system setup
require(tidyverse)
require(azmpdata)
require(marea)
require(gratia)
require(gradientForest)
require(mgcv)
require(here)
```

# gather indicators and pressures

```{r}

data("eco_indicators")


#separate by region
# indicators used in Irvine et al. 2025 reasoning was that they didn't use any indicators calculated that had landings (not available at set level), MargalefRichness_ALL, Heips_ALL, LargeFishIndicator, MeanLengthAbundance, Biomass_Piscivore, CCondition_Finfish, CCondition_MBenthivore, CCondition_Zoopiscivore, CCondition_Planktivores, CCondition_LBenthivore, CVBiomass, MeanLifespan, BiomassTL2, Biomass_Skates, Biomass_All, Biomass_Flatfish, Biomass_Gadoids, Biomass_invertebrates

nafo_ei_4VN<-eco_indicators[eco_indicators@data$region=="4VN"]
nafo_ei_4VS<-eco_indicators[eco_indicators@data$region=="4VS"]
nafo_ei_4W<-eco_indicators[eco_indicators@data$region=="4W"]
nafo_ei_4X<-eco_indicators[eco_indicators@data$region=="4X"]
nafo_ind_4X<-nafo_ei_4X@data |> 
    select(margalefrichness=MargalefRichness_ALL_value,
           heips=Heips_ALL_value,
           lfi=LargeFishIndicator_value,
           mla=MeanLengthABUNDANCE_value,
           biomasspiscivore=BIOMASS_PISCIVORE_value
           )


ESS_ei<-eco_indicators[eco_indicators@data$region=="ESS"]
WSS_ei<-eco_indicators[eco_indicators@data$region=="WSS"]

# pull environmental indicators and fishing from azmpdata?
data("azmp_bottom_temperature")
# nafo level, SST,Bottom Temp,bottom temperature anomalies, sst_anomalies,  Cal fin, zooplankton, Integrated chl, Nitrate (stations, lines?)
nafo_bt_4VN<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="4VN"]
nafo_bt_4VS<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="4VS"]
nafo_bt_4W<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="4W"]
nafo_bt_4X<-eco_indicators[eco_indicators@data$region=="4X"]

nafo_pres_4X<-nafo_ei_4X@data |> 
    select(landall=landings_ALL_value, 
           landforage=landings_FORAGE.L_value,
           fpall=FP_ALL_value,
           fpgadoids=FP_GADOIDS_value #no spaces or underscores!
           )
#remove weird 2021 fpall value >3
nafo_pres_4X[52,"fpall"]<-NA

nafo_sst_4X<-azmpdata::Derived_Annual_Broadscale |>
        filter(area=="4XSS")|>
        select(YEAR=year,SST=sea_surface_temperature_from_satellite)

nafo_land_4X<-eco_indicators[eco_indicators@data$region=="4X"]
  
# Shelf level indicators, NAO_w, AMO, Bloom stats
  ESS_bt<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="ESS"]
  WSS_bt<-azmp_bottom_temperature[azmp_bottom_temperature@data$region=="WSS"]
  nao<-data("nao")
  amo<-data("amo")
  
```

# Step 1. run GAM and GLMs for each indicator x pressure combo

```{r}
library(mgcv)
library(gratia)

# Identify indicator and pressure tables
indicators <- nafo_ind_4X
pressures <- nafo_pres_4X

# Assume nafo_ind_4X and nafo_pres_4X are loaded tables with matching row order

model_list <- list()

for (a_col in names(indicators)) {       # Indicator variables
  for (b_col in names(pressures)) {    # Pressure variables
    tryCatch({
      # Combine indicator and pressure into a single dataframe with correct col names
      df <- data.frame(
        resp = indicators[[a_col]],
        pres = pressures[[b_col]]
      )
      colnames(df) <- c(a_col, b_col)
      df <- na.omit(df)
      
      # Only fit if sufficient data and nonzero variance
      if (nrow(df) < 10 || var(df[[b_col]]) == 0 || var(df[[a_col]]) == 0) next
      
      # Fit GAM with real variable names
      gam_formula <- as.formula(paste(a_col, "~ s(", b_col, ")"))
      gam_model <- gam(gam_formula, data = df, method = "REML", family = tw())

      # Fit GLM with same variables
      glm_formula <- as.formula(paste(a_col, "~", b_col))
      glm_model <- gam(glm_formula, data = df, method = "REML", family = tw())

      # Store models with unique names for later reference
      model_name_base <- paste(a_col, b_col, sep = "_vs_")
      model_list[[paste0(model_name_base, "_GAM")]] <- gam_model
      model_list[[paste0(model_name_base, "_GLM")]] <- glm_model
      
    }, error = function(e) {
      message("Error fitting models for ", a_col, " vs ", b_col, ": ", e$message)
    })
  }
}
```

# Step 2. Pull model results and create a data table of model results

```{r}
# Initialize comparison results data frame
comparison_results <- data.frame()

# Loop through model_list to extract diagnostics
for (model_name in names(model_list)) {
  tryCatch({
    model <- model_list[[model_name]]
    model_summary <- summary(model)
    
    # Determine model type and variable names
    parts <- strsplit(model_name, "_vs_")[[1]]
    response <- parts[1]
    predictor_and_type <- parts[2]
    
    # Separate predictor and model type
    predictor <- sub("_.*$", "", predictor_and_type)
    model_type <- sub("^.*_", "", predictor_and_type)
    
    # Extract diagnostics
    aic_val <- AIC(model)
    r_squared <- model_summary$r.sq
    edf <- if ("edf" %in% names(model_summary)) model_summary$edf[1] else NA
    gcv_score <- model$gcv.ubre
    
    # Extract p-value depending on model type
p_value <- tryCatch({
  if (model_type == "GAM") {
    model_summary$s.table[1, "p-value"]
  } else if (model_type == "GLM") {
    # Use predictor name directly
    if (predictor %in% rownames(model_summary$p.table)) {
      model_summary$p.table[predictor, "Pr(>|t|)"]
    } else {
      # Fallback: assume 2nd row is predictor
      model_summary$p.table[2, "Pr(>|t|)"]
    }
  } else {
    NA
  }
}, error = function(e) NA)
    # Store results
    comparison_results <- rbind(comparison_results, data.frame(
      Response = response,
      Predictor = predictor,
      Model_Type = model_type,
      AIC = aic_val,
      R_squared = r_squared,
      EDF = edf,
      GCV_UBRE = gcv_score,
      p_value = p_value,
      stringsAsFactors = FALSE
    ))
    
  }, error = function(e) {
    message("Error extracting diagnostics for ", model_name, ": ", e$message)
  })
}

# View comparison table
print(comparison_results)
```

# Step 3. Model selection proceedure selecting lower GCV scores from the comparison_results table

```{r}
library(dplyr)

# Compare GAM vs GLM for each Response-Predictor pair
model_selection <- comparison_results %>%
  group_by(Response, Predictor) %>%
  filter(Model_Type %in% c("GAM", "GLM")) %>%
  arrange(GCV_UBRE) %>%
  slice(1)  # Select model with lowest GCV


print(model_selection)

```

# Step 4. identify robust treshold crossings
```{r}
crossings_list <- list()

for (model_name in names(model_list)) {
  if (grepl("_GAM$", model_name)) {
    model <- model_list[[model_name]]
    # Extract predictor name
    parts <- strsplit(model_name, "_vs_")[[1]]
    response <- parts[1]
    predictor <- sub("_GAM$", "", parts[2])
    term_name <- paste0("s(", predictor, ")")
    deriv <- derivatives(model, term = term_name, partial_match = TRUE)
    dvals <- deriv$.derivative
    xvals <- deriv[[predictor]]
    sign_change_idx <- which(diff(sign(dvals)) != 0)
    crossings <- numeric(0)
    for (i in seq_along(sign_change_idx)) {
      idx <- sign_change_idx[i]
      x1 <- xvals[idx]
      x2 <- xvals[idx + 1]
      y1 <- dvals[idx]
      y2 <- dvals[idx + 1]
      crossing <- x1 - y1 * (x2 - x1) / (y2 - y1)
      crossings <- c(crossings, crossing)
    }
    crossings_list[[model_name]] <- crossings
  }
}

# Only keep models with at least one crossing
crossings_df <- do.call(rbind, lapply(names(crossings_list), function(name) {
  crossings <- crossings_list[[name]]
  if (length(crossings) > 0) {
    data.frame(
      Model = name,
      Crossing_Index = seq_along(crossings),
      Predictor_Zero_Crossing = crossings
    )
  } else {
    NULL
  }
}))

print(crossings_df)
```

# Step 5. Create plots of significant GLM and GAMs showing significant deriviative crossings in the GAM plots with a red line at the x intercept

```{r}
library(ggplot2)
library(patchwork)

plot_list <- list()

for (i in seq_len(nrow(model_selection))) {
  row <- model_selection[i, ]
  response <- row$Response
  predictor <- row$Predictor
  model_type <- row$Model_Type
  full_model_name <- paste0(response, "_vs_", predictor, "_", model_type)
  model <- model_list[[full_model_name]]

  if (!is.null(model)) {
    if (model_type == "GAM") {
  # Gather threshold crossings for this model, if any
  crossings <- crossings_df$Predictor_Zero_Crossing[crossings_df$Model == full_model_name]
  
  # Build prediction data with the correct column name!
  pred_vals <- seq(
    min(model$model[[predictor]], na.rm = TRUE), 
    max(model$model[[predictor]], na.rm = TRUE), 
    length.out = 200
  )
  pred_df <- setNames(data.frame(pred_vals), predictor)  # <-- THE FIX

  # Predict
  pred_df$fit <- predict(model, newdata = pred_df, type = "response")
  pred_conf <- predict(model, newdata = pred_df, type = "link", se.fit = TRUE)
  pred_df$se <- pred_conf$se.fit
  # Compute approximate confidence interval on response scale
  pred_df$upper <- model$family$linkinv(pred_conf$fit + 1.96 * pred_df$se)
  pred_df$lower <- model$family$linkinv(pred_conf$fit - 1.96 * pred_df$se)

  # Plot GAMs
  p <- ggplot() +
    geom_point(data = model$model, aes_string(x = predictor, y = response), alpha = 0.5) +
    geom_line(data = pred_df, aes_string(x = predictor, y = "fit"), color = "black", size = 1) +
    geom_ribbon(data = pred_df, aes_string(x = predictor, ymin = "lower", ymax = "upper"), fill = "black", alpha = 0.2) +
    labs(
      # title = paste("GAM:", response, "vs", predictor),
      x = predictor,
      y = response
    ) +
    theme_minimal()

      # Add crossings as vertical lines (if present)
      if (length(crossings) > 0) {
        for (v in crossings) {
          p <- p + geom_vline(xintercept = v, linetype = "dashed", color = "red")
        }
      }
      
    } else if (model_type == "GLM") {
      df <- model$model
      df$fit <- fitted(model)
      df$se <- sqrt(model$scale) * predict(model, type = "link", se.fit = TRUE)$se.fit
      df$upper <- df$fit + df$se
      df$lower <- df$fit - df$se
# Plot GLMs
      p <- ggplot(df, aes_string(x = predictor, y = response)) +
        geom_point(alpha = 0.6) +
        geom_line(aes(y = fit), color = "blue", size = 1) +
        geom_ribbon(aes(ymin = lower, ymax = upper), fill = "blue", alpha = 0.2) +
        labs(
          # title = paste("GLM:", response, "vs", predictor),
          x = predictor,
          y = response
        ) +
        theme_minimal()
    }
    plot_list[[full_model_name]] <- p
  }
}

# Combine plots (will plot up to 4 per row, adjust ncol as desired)
combined_plot <- wrap_plots(plot_list, ncol = 4)
print(combined_plot)

```

# Step 6. Save outputs

```{r}

# Ensure output folder exists
dir.create(here::here("results"), showWarnings = FALSE, recursive = TRUE)
dir.create(here::here("figures"), showWarnings = FALSE, recursive = TRUE)

# Set region
region <- "4X"

# Save to results folder
save(
  plot_list,
  combined_plot,
  model_list,
  comparison_results,
  crossings_df,
  file = here("results", paste0("nafo_outputs_", region, ".RData"))
)


# File name (e.g., figures/nafo_combined_<region>.png)
png_path <- here::here("figures", paste0("nafo_combined_", region, ".png"))

# Save combined plot (adjust width/height as needed)
ggsave(
  filename = png_path,
  plot = combined_plot,
  width = 16, height = 12, units = "in", dpi = 300, bg = "white"
)


```


