---
title: "Untitled"
author: "Jamie"
date: "2024-04-30"
output: html_document
---

This document consolidates the steps from SSdata to marinidicators for caluclating ecological indicators from the Maritimes Region.


```{r}
#setup
library(devtools)
#library(SSdata)
library(here)
library(tidyverse)
library(readxl)
library(marindicators)
library(missForest)
install_github("jamiectam/SSdata")


```

Setup parameters
```{r}
path<-file.path(here())
start.year <- 2022 #recommend just updating one or 2 years to save time
end.year <- 2022
areas.RV <- c("esswss", "shelf", "nafo", "strat")
areas.land <- c("esswss", "shelf", "nafo")

```


This starts after running SSdata::biomassData()
                          SSdata::extractLAND()
which requires username and password for accessing databases   


## Update biomass and landings data
```{r}
#extracts RV survey data, requires passwords 
biomassData(path=path, 
            s.strat = 440, 
            e.strat = 495, 
            s.year=start.year, 
            e.year=end.year, 
            vessel.correction = TRUE)

#resulting Rdata files should go into the correct folder (have not yet tested) 

extractLAND(path=path,
              e.year = end.year)

```


## Calculate dataframes for specific areas                       
```{r}
#can also use update_LAND=TRUE

LANDdataframe(path=path, areas = c("shelf", "esswss", "nafo"), update_LAND=FALSE,
           e.year=end.year, csv = TRUE, rdata = TRUE)

stratifyBiomass(path=path, s.year=start.year, e.year=end.year, areas=areas.RV,
                lengthbased=TRUE, 
                 qadjusted=TRUE)

stratifyBiomass(path=path, s.year=start.year, e.year=end.year, areas=areas.RV,
                lengthbased=TRUE, 
                 qadjusted=FALSE)

stratifyBiomass(path=path, s.year=start.year, e.year=end.year, areas=areas.RV,
                lengthbased=FALSE, 
                 qadjusted=TRUE)

stratifyBiomass(path=path, s.year=start.year, e.year=end.year, areas=areas.RV,
                lengthbased=FALSE, 
                 qadjusted=FALSE)

 RVdataframe(path = path, s.year = start.year, e.year = end.year, areas = areas.RV,
            lengthbased = TRUE,
            qadjusted = TRUE,
            update_RV = FALSE,
            csv = TRUE, rdata = TRUE)

RVdataframe(path = path, s.year = start.year, e.year = end.year, areas = areas.RV,
            lengthbased = TRUE,
            qadjusted = FALSE,
            update_RV = FALSE,
            csv = TRUE, rdata = TRUE)

RVdataframe(path = path, s.year = start.year, e.year = end.year, areas = areas.RV,
            lengthbased = FALSE,
            qadjusted = TRUE,
            update_RV = FALSE,
            csv = TRUE, rdata = TRUE)

RVdataframe(path = path, s.year = start.year, e.year = end.year, areas = areas.RV,
            lengthbased = FALSE,
            qadjusted = FALSE,
            update_RV = FALSE,
            csv = TRUE, rdata = TRUE)

LWdataframe(path=path, s.year=start.year, e.year=end.year, areas = c("shelf", "esswss", "nafo","strat"), update_LW = FALSE, csv = TRUE, rdata = TRUE)

```
                        
                        
## Update herring                          
```{r}
path <- file.path(here())

source(paste(path, "/Extra Info/convert_herring_length.R", sep = ""))

length.cutoff <- 23

# IMPORT DATA -------------------------------------------------------------

# adjusted-VPA biomass index 
adj_dat <- read.csv(paste(path, "/output/RV/herring_data/herring_biomass_adjusted.csv", sep = "")) 
adj_VPA <- adj_dat %>% 
  filter(YEAR >= 1970, YEAR < 1999) %>%  # filter for years >= 1970 & years < 1999
  mutate(BIOMASS = BIOMASS * 1000) %>%   # convert from tonnes to kg
  rename("BIOMASS_INDEX" = "BIOMASS")    # rename biomass column to differentiate from RV BIOMASS

# acoustic abundance at length: numbers (millions) at length (cm)
acoustic.abund <- read_excel(paste(
  path, "/output/RV/herring_data/Acoustic Number and Biomass at Length_200313.xlsx", sep = ""), sheet = 2)
acoustic.abund.tidy <- pivot_longer(acoustic.abund, 
                                    cols = c(2:ncol(acoustic.abund)), 
                                    names_to = "LENGTH", values_to = "ABUNDANCE") %>% 
  rename("YEAR" = "Year") %>% 
  mutate(ABUNDANCE = ABUNDANCE * 10^6, LENGTH = as.numeric(LENGTH))  # convert to numbers

# acoustic biomass at length: weight (metric tonnes) at length (cm)
acoustic.bio <- read_excel(paste(
  path, "/output/RV/herring_data/Acoustic Number and Biomass at Length_200313.xlsx", sep = ""), sheet = 3)
acoustic.bio.tidy <- pivot_longer(acoustic.bio, 
                                  cols = c(2:ncol(acoustic.bio)), 
                                  names_to = "LENGTH", values_to = "BIOMASS") %>% 
  rename("YEAR" = "Year") %>% 
  mutate(BIOMASS = BIOMASS * 10^3, LENGTH = as.numeric(LENGTH)) # convert to weight in kg

# Join biomass and abundance data
acoustic <- full_join(acoustic.bio.tidy, acoustic.abund.tidy, by = c("YEAR", "LENGTH")) %>% 
  filter(LENGTH >= length.cutoff)  # subset for lengths >= 23 cm

###########################################################################
# WSS Scale ---------------------------------------------------------------
###########################################################################

# Import and tidy stratified RV data at the WSS scale
RV_wss <- read.csv(paste(path, "/output/RV/esswss/esswss_lengthbased_notqadj.csv", sep = "")) %>% 
  filter(SPECIES == 60,                # subset for herring
         ID == "WSS",                  # subset for WSS
         YEAR >= 1970, 
         YEAR < 1999) %>%              # subset for years < 1999 (acoustic data used after 1999)
  convert_herring_length() %>%         # convert all lengths to total length in cm
  select(-ID, -SPECIES) %>%
  filter(LENGTH >= length.cutoff)      # subset for lengths >= 23 cm (mature fish)

###############
## Aggregate ##
###############

### Before 1999

# Calculate annual mean weight (kg) from RV data
MEAN_WT <- RV_wss %>% 
  group_by(YEAR) %>% 
  summarise(BIOMASS = sum(BIOMASS), ABUNDANCE = sum(ABUNDANCE)) %>% # total biomass and abundance each year
  mutate(MEAN_WT = BIOMASS/ABUNDANCE) %>%                           # mean weight each year
  select(-BIOMASS, -ABUNDANCE)

# Merge adjusted-VPA index with annual mean weight and calculate abundance
adj_agg <- full_join(adj_VPA, MEAN_WT, by = "YEAR") %>% 
  mutate(ABUNDANCE = BIOMASS_INDEX / MEAN_WT) %>% 
  select(-MEAN_WT) %>% 
  rename("BIOMASS" = "BIOMASS_INDEX")

rm(MEAN_WT)

### After 1999

# Sum biomass and abundance over all length classes each year
acoustic_agg <- acoustic %>%  
  group_by(YEAR) %>% 
  summarize(BIOMASS = sum(BIOMASS), ABUNDANCE = sum(ABUNDANCE))

# Total biomass and abundance indices for WSS:
HER_agg_wss <- rbind(adj_agg, acoustic_agg)

###############
## Length #####
###############

### Before 1999:
  
# Calculate mean weight (kg) at length (cm) from RV data 
MEAN_WT_length <- RV_wss %>% mutate(MEAN_WT = BIOMASS/ABUNDANCE) %>% 
  select(-BIOMASS, -ABUNDANCE)

# Calculate propotion of total biomass at each length from RV data;
# Allocate the biomass index into length classes based on RV proportions;
# Calculate abundance from mean weight
adj_length <- RV_wss %>%  select(-ABUNDANCE) %>% 
  group_by(YEAR) %>% 
  mutate(TOTAL_BIOMASS = sum(BIOMASS))%>%            # total biomass each year
  ungroup() %>% 
  mutate(PROP_BIOMASS = BIOMASS/TOTAL_BIOMASS) %>%   # proportion of total biomass at each length
  full_join(adj_VPA, by = "YEAR") %>%                # merge with adjusted-VPA index        
  mutate(BIOMASS_INDEX_LENGTH = PROP_BIOMASS * BIOMASS_INDEX) %>%  # allocate adjusted-VPA index into length classes
  select(YEAR, LENGTH, BIOMASS_INDEX_LENGTH) %>% 
  left_join(MEAN_WT_length, by = c("YEAR", "LENGTH")) %>%          # merge with mean weight dataframe
  mutate(ABUNDANCE_INDEX_LENGTH = BIOMASS_INDEX_LENGTH / MEAN_WT) %>%   # calculate abundance
  rename("BIOMASS" ="BIOMASS_INDEX_LENGTH", "ABUNDANCE" = "ABUNDANCE_INDEX_LENGTH") %>% 
  select(-MEAN_WT) 

rm(MEAN_WT_length)

### After 1999:
  
# Acoustic data

# Biomass and abundance indices at length for WSS:
HER_length_wss <- rbind(adj_length, acoustic)

###########################################################################
# Strata Scale ------------------------------------------------------------
###########################################################################

# Import and tidy stratified RV data at the strata scale. 
RV_strat <- read.csv(paste(path, "/output/RV/strat/strat_lengthbased_notqadj.csv", sep = "")) %>% 
  filter(SPECIES == 60,                  # subset for herring
         ID >= 470,                      # subset for strata on the WSS
         YEAR >= 1970,                 
         YEAR <= 2022) %>% 
  convert_herring_length() %>%           # convert all lengths to total length in cm
  filter(LENGTH >= length.cutoff) %>%    # subset for lengths >= 23 cm
  select(-SPECIES)

# Calculate proportion of total biomass and abundance in each strata from RV data
RV_prop <- RV_strat %>%  
  group_by(YEAR, ID) %>% 
  summarise(BIOMASS = sum(BIOMASS), ABUNDANCE = sum(ABUNDANCE))%>%           # total biomass in each strata for each year
  ungroup() %>% 
  group_by(YEAR) %>%            
  mutate(TOTAL_BIOMASS = sum(BIOMASS), TOTAL_ABUNDANCE = sum(ABUNDANCE)) %>% # total biomass each year
  ungroup() %>% 
  # proportion of total biomass and abundance in each strata
  mutate(PROP_BIOMASS = BIOMASS/TOTAL_BIOMASS, PROP_ABUNDANCE = ABUNDANCE/TOTAL_ABUNDANCE) %>% 
  select(YEAR, ID, PROP_BIOMASS, PROP_ABUNDANCE)

###############
## Aggregate ##
###############

### Before 1999:
  
# Calculate annual mean weight (kg) for each strata from RV data 
MEAN_WT <- RV_strat %>% filter(YEAR < 1999) %>% 
  group_by(YEAR, ID) %>% 
  summarise(BIOMASS = sum(BIOMASS), ABUNDANCE = sum(ABUNDANCE)) %>% 
  mutate(MEAN_WT = BIOMASS/ABUNDANCE) %>% 
  select(-BIOMASS, -ABUNDANCE) %>%  
  ungroup() 

# Allocate the adjusted biomass index into each strata and then calculate abundance
adj_agg_strat <- inner_join(RV_prop, adj_VPA, by = "YEAR") %>%          # merge RV proportions with adjusted-VPA data
  mutate(BIOMASS_INDEX_STRAT = PROP_BIOMASS * BIOMASS_INDEX) %>%        # allocate biomass index into each strata
  select(YEAR, ID, BIOMASS_INDEX_STRAT) %>% 
  inner_join(MEAN_WT, by = c("YEAR", "ID")) %>%                         #  merge with mean weight dataframe
  mutate(ABUNDANCE_INDEX_STRAT = BIOMASS_INDEX_STRAT / MEAN_WT) %>%     # calculate abundance index
  rename("BIOMASS" ="BIOMASS_INDEX_STRAT", "ABUNDANCE" = "ABUNDANCE_INDEX_STRAT") %>% 
  select(-MEAN_WT) 

### After 1999:
  
# Sum biomass and abundance over all length classes and then allocate biomass and abundance into strata
acoustic_agg_strat <- acoustic %>%  
  group_by(YEAR) %>% 
  summarize(BIOMASS = sum(BIOMASS), ABUNDANCE = sum(ABUNDANCE)) %>%  # total biomass and abundance each year
  inner_join(RV_prop, by = "YEAR") %>%                                # merge RV proportions with acoustic data
  # allocate biomass and abundance into strata based on RV proportions
  mutate(BIOMASS = BIOMASS * PROP_BIOMASS, ABUNDANCE = ABUNDANCE * PROP_ABUNDANCE) %>% 
  select(YEAR, ID, BIOMASS, ABUNDANCE)

# Total biomass and abundance indices at for each strata on WSS:
HER_agg_strat <- rbind(adj_agg_strat, acoustic_agg_strat)

###############
## Length #####
###############

### Before 1999:
  
# Calculate proportion of RV biomass and abundance in each strata and length class
RV_prop_length <- RV_strat %>% filter(YEAR < 1999) %>% 
  group_by(YEAR) %>% 
  # total biomass & abundance over all strata & length classes
  mutate(TOTAL_BIOMASS = sum(BIOMASS), TOTAL_ABUNDANCE = sum(ABUNDANCE)) %>%                   
  ungroup() %>% 
  # proportion of total biomass and abundance in each strata and length class
  mutate(PROP_BIOMASS = BIOMASS/TOTAL_BIOMASS, PROP_ABUNDANCE = ABUNDANCE/TOTAL_ABUNDANCE) %>% 
  select(YEAR, ID, LENGTH, PROP_BIOMASS, PROP_ABUNDANCE)

# Calculate mean weight (kg) at length (cm) from RV data 
MEAN_WT_length <- RV_strat %>%  filter(YEAR < 1999) %>% 
  mutate(MEAN_WT = BIOMASS/ABUNDANCE) %>% 
  select(-BIOMASS, -ABUNDANCE)

# Allocate the adjusted-VPA biomass index into each strata and length class and then calculate abundance
adj_length <-  inner_join(RV_prop_length, adj_VPA, by = "YEAR") %>%     
  # allocate biomass index into each strata & length class based on RV propotions
  mutate(BIOMASS_INDEX_STRAT = PROP_BIOMASS * BIOMASS_INDEX) %>%      
  select(YEAR, ID, LENGTH, BIOMASS_INDEX_STRAT) %>% 
  inner_join(MEAN_WT_length, by = c("YEAR", "ID", "LENGTH")) %>%     # merge with mean weight dataframe
  mutate(ABUNDANCE_INDEX_STRAT = BIOMASS_INDEX_STRAT / MEAN_WT) %>%  # calculate abundance
  rename("BIOMASS" ="BIOMASS_INDEX_STRAT", "ABUNDANCE" = "ABUNDANCE_INDEX_STRAT") %>% 
  select(-MEAN_WT) 

### After 1999:

# Use RV data to calculate proportion of biomass and abundance from each length class in each strata. 
## (Note: this is NOT the same as in Before 1999. Before 1999, we allocate the total biomass into strata and length class. 
## Here, we allocate the biomass in a given length class into each strata.)
RV_length_acoustic <- RV_strat %>%  
  group_by(YEAR, LENGTH) %>% 
  # total biomass and abundance in each length class (sum over all strata)
  mutate(BIOMASS_LENGTH = sum(BIOMASS), ABUNDANCE_LENGTH = sum(ABUNDANCE)) %>% 
  ungroup() %>% 
  # proportion of biomass and abundance from each length class in each strata
  mutate(PROP_BIOMASS = BIOMASS/BIOMASS_LENGTH, PROP_ABUNDANCE = ABUNDANCE/ABUNDANCE_LENGTH) %>% 
  select(YEAR, ID, LENGTH, PROP_BIOMASS, PROP_ABUNDANCE)

# Allocate the biomass and abundance at each length into strata
acoustic_length <- acoustic %>% 
  inner_join(RV_length_acoustic, by = c("YEAR", "LENGTH"))%>% 
  mutate(BIOMASS = BIOMASS * PROP_BIOMASS, ABUNDANCE = ABUNDANCE * PROP_ABUNDANCE) %>% 
  select(YEAR, ID, LENGTH, BIOMASS, ABUNDANCE)

# Biomass and abundance indices at length for each strata on WSS:
HER_length_strat <- rbind(adj_length, acoustic_length)



# Export ------------------------------------------------------------------

dir.create(paste(path, "/output/RV/herring_output", sep = ""), recursive = FALSE, showWarnings = T)

write_csv(HER_agg_wss, 
          path = paste(path, "/output/RV/herring_output/wss_notlengthbased_herring.csv", sep = ""))

write_csv(HER_length_wss, 
          path = paste(path, "/output/RV/herring_output/wss_lengthbased_herring.csv", sep = ""))

write_csv(HER_agg_strat, 
          path = paste(path, "/output/RV/herring_output/strat_notlengthbased_herring.csv", sep = ""))

write_csv(HER_length_strat, 
          path = paste(path, "/output/RV/herring_output/strat_lengthbased_herring.csv", sep = ""))

```
  
## Merge RV and herring
```{r}
path <- file.path(here())

source(paste(path, "/Extra Info/convert_herring_length.R", sep = ""))

#  WSS AGG ---------------------------------------------------------

# herring index: add ID and SPECIES columns
HER_wss_agg <- read_csv(paste(path, "/output/RV/herring_output/wss_notlengthbased_herring.csv", sep = "")) %>% 
  mutate(ID = "WSS", SPECIES = 60)

# RV data
esswss_agg <- read_csv(paste(path, "/output/RV/esswss/esswss_notlengthbased_qadj.csv", sep = ""))

# ess: no changes
ess_agg <- esswss_agg %>% filter(ID == "ESS")

# wss: remove RV herring observations and add index
wss_agg <- esswss_agg %>% filter(ID == "WSS", SPECIES != 60) %>% 
  rbind(HER_wss_agg)

# put ESS and WSS back together
esswss_agg_OUT <- rbind(ess_agg, wss_agg)


#  WSS LENGTH ---------------------------------------------------------

# herring index: add ID and SPECIES columns
HER_wss_length <- read_csv(paste(path, "/output/RV/herring_output/wss_lengthbased_herring.csv", sep = ""))  %>% 
  mutate(ID = "WSS", SPECIES = 60)

# RV data
esswss_length <-  read_csv(paste(path, "/output/RV/esswss/esswss_lengthbased_qadj.csv", sep = ""))

# ess: convert herring lengths to total length in cm
ess_length <- esswss_length %>% filter(ID == "ESS") %>% 
  convert_herring_length()

# wss: remove RV herring observations and add index
wss_length <- esswss_length %>% filter(ID == "WSS", SPECIES != 60) %>% 
  rbind(HER_wss_length)

# put ESS and WSS back together
esswss_length_OUT <- rbind(ess_length, wss_length)


#  NAFO AGG ---------------------------------------------------------

# herring index: 4X is the same as WSS
HER_4X_agg <- read_csv(paste(path, "/output/RV/herring_output/wss_notlengthbased_herring.csv", sep = "")) %>% 
  mutate(ID = "4X", SPECIES = 60)

# RV data
nafo_agg <- read_csv(paste(path, "/output/RV/nafo/nafo_notlengthbased_qadj.csv", sep = ""))

# 4VN, 4VS, 4W: no changes
VW_agg <- nafo_agg %>% filter(ID != "4X")

# 4X: remove RV herring observations and add index
X_agg <- nafo_agg %>% filter(ID == "4X", SPECIES != 60) %>% 
  rbind(HER_4X_agg)

# put ESS and WSS back together
nafo_agg_OUT <- rbind(VW_agg, X_agg)


#  NAFO LENGTH ---------------------------------------------------------

# herring index: add ID and SPECIES columns
HER_4X_length <- read_csv(paste(path, "/output/RV/herring_output/wss_lengthbased_herring.csv", sep = ""))  %>% 
  mutate(ID = "4X", SPECIES = 60)

# RV data
nafo_length <-  read_csv(paste(path, "/output/RV/nafo/nafo_lengthbased_qadj.csv", sep = ""))

# 4VN, 4VS, 4W: convert herring lengths to total length in cm
VW_length <- nafo_length %>% filter(ID != "4X") %>% 
  convert_herring_length()

# wss: remove RV herring observations and add index
X_length <- nafo_length %>% filter(ID == "4X", SPECIES != 60) %>% 
  rbind(HER_4X_length)

# put ESS and WSS back together
nafo_length_OUT <- rbind(VW_length, X_length)


#  strat AGG ---------------------------------------------------------

# herring index: add SPECIES columns
HER_strat_agg <- read_csv(paste(path, "/output/RV/herring_output/strat_notlengthbased_herring.csv", sep = "")) %>% 
  mutate(SPECIES = 60)

# RV data
strat_agg <- read_csv(paste(path, "/output/RV/strat/strat_notlengthbased_qadj.csv", sep = ""))

# ess: no changes
ess_strat_agg <- strat_agg %>% filter(ID < 470)

# strat: remove RV herring observations and add index
wss_strat_agg <- strat_agg %>% filter(ID >= 470, SPECIES != 60) %>% 
  rbind(HER_strat_agg)

# put ESS and strat back together
strat_agg_OUT <- rbind(ess_strat_agg, wss_strat_agg)


#  strat LENGTH ---------------------------------------------------------

# herring index: add SPECIES columns
HER_strat_length <- read_csv(paste(path, "/output/RV/herring_output/strat_lengthbased_herring.csv", sep = "")) %>% 
  mutate(SPECIES = 60)

# RV data
strat_length <- read_csv(paste(path, "/output/RV/strat/strat_lengthbased_qadj.csv", sep = ""))

# ess: convert herring lengths to total length in cm
ess_strat_length <- strat_length %>% filter(ID < 470) %>% 
   convert_herring_length()

# strat: remove RV herring observations and add index
wss_strat_length <- strat_length %>% filter(ID >= 470, SPECIES != 60) %>% 
  rbind(HER_strat_length)

# put ESS and strat back together
strat_length_OUT <- rbind(ess_strat_length, wss_strat_length)



# Export ------------------------------------------------------------------

dir.create(paste(path, "/final_output", sep = ""), recursive = FALSE, showWarnings = FALSE)

write_csv(esswss_agg_OUT,
          path = paste(path, "/final_output/esswss_notlengthbased_qadj_final.csv", sep = ""))

write_csv(esswss_length_OUT,
          path = paste(path, "/final_output/esswss_lengthbased_qadj_final.csv", sep = ""))

write_csv(nafo_agg_OUT,
          path = paste(path, "/final_output/nafo_notlengthbased_qadj_final.csv", sep = ""))

write_csv(nafo_length_OUT,
          path = paste(path, "/final_output/nafo_lengthbased_qadj_final.csv", sep = ""))

write_csv(strat_agg_OUT,
          path = paste(path, "/final_output/strat_notlengthbased_qadj_final.csv", sep = ""))

write_csv(strat_length_OUT,
          path = paste(path, "/final_output/strat_lengthbased_qadj_final.csv", sep = ""))


# LENGTH-WEIGHT ---------------------------------------------------

# import LW data from SSdata output
## and convert herring lengths to total length in cm
lw_esswss <-  read_csv(paste(path, "/output/LengthWeight/esswss_LengthWeight.csv", sep = "")) %>% 
 convert_herring_length() %>% 
  arrange(ID, YEAR, SPECIES, LENGTH)

# import LW data from SSdata output
## and convert herring lengths to total length in cm
lw_nafo <-  read_csv(paste(path, "/output/LengthWeight/nafo_LengthWeight.csv", sep = "")) %>% 
  convert_herring_length() %>% 
  arrange(ID, YEAR, SPECIES, LENGTH)

# import LW data from SSdata output
## and convert herring lengths to total length in cm
lw_strat <-  read_csv(paste(path, "/output/LengthWeight/strat_LengthWeight.csv", sep = "")) %>% 
  convert_herring_length() %>% 
  arrange(ID, YEAR, SPECIES, LENGTH)

# Export ------------------------------------------------------------------

write_csv(lw_esswss, 
          path = paste(path, "/final_output/esswss_LengthWeight.csv", sep = ""))

write_csv(lw_nafo, 
          path = paste(path, "/final_output/nafo_LengthWeight.csv", sep = ""))

write_csv(lw_strat, 
          path = paste(path, "/final_output/strat_LengthWeight.csv", sep = ""))


```
      
      
## calculate NAFO indicators                          
```{r}
path <- file.path(here())

source(paste(path, "/Extra Info/get_NA_index_add_NAs.R", sep = ""))

dir.create(paste(path, "/INDICATORS", sep = ""), recursive = FALSE, showWarnings = FALSE)
export.filepath <- file.path(paste(path, "/INDICATORS", sep  = ""))

start_year <- 1970
end_year <- 2022
ind.years <- c(start_year:end_year)

# Extra data
species.groups <-  read.csv(paste(path, "/Extra Info/SpeciesGroups.csv", sep = ""), header = TRUE, sep = ",")
species.info <-  read.csv(paste(path, "/Extra Info/SpeciesInfo.csv", sep = ""), header = TRUE, sep = ",")

# Group names
ratio.groups <- data.frame(rbind(c("PELAGIC", "GROUNDFISH"), c("PREDATORS", "ALL")))
names(ratio.groups) <- c("group1", "group2")

trophicguild.groups <- c("LBENTHIVORE", "MBENTHIVORE", "PISCIVORE", "PLANKTIVORE", "ZOOPISCIVORE")

condition.groups <- c("FINFISH", "LBENTHIVORE", "MBENTHIVORE", "PISCIVORE", "PLANKTIVORE", "ZOOPISCIVORE")

# resource potential
resource.groups <- c("ALL", "CLUPEIDS", "FINFISH", "FLATFISH", "FORAGE", 
                     "GADOIDS", "GROUNDFISH", "PELAGIC", "SKATES")

# allIndicators
minTL <- c(0, 3.25)
landings.groups <- c("ALL", "CLUPEIDS.L", "FINFISH.L", "FLATFISH.L", "FORAGE.L",
                     "GADOIDS.L", "GROUNDFISH.L", "INVERTEBRATES.L","LARGE_PELAGIC.L",  "SKATES.L")
FP.groups <- data.frame(rbind(c("ALL", "ALL"), 
                              c("CLUPEIDS", "CLUPEIDS.L"),
                              c("FINFISH", "FINFISH.L"),
                              c("FLATFISH", "FLATFISH.L"),
                              c("FORAGE", "FORAGE.L"),
                              c("GADOIDS", "GADOIDS.L"),
                              c("GROUNDFISH", "GROUNDFISH.L"),
                              c("SKATES", "SKATES.L")))
names(FP.groups) <- c("group.X", "group.land")

# NAFO Data ------------------------------------------------------------------

area <- "nafo"

# fishery-independent data
RV <- read.csv(paste(path, "/final_output/nafo_notlengthbased_qadj_final.csv", sep = ""), 
               head = TRUE, sep = ",", stringsAsFactors = FALSE)
RV_length <- read.csv(paste(path, "/final_output/nafo_lengthbased_qadj_final.csv", sep = ""), 
                      head = TRUE, sep = ",", stringsAsFactors = FALSE)

# commercial landings data
# landings converted from tonnes to kg to be on the same scale as the RV data (required for the FP indicators)
land <- read.csv(paste(path, "/output/Landings/nafo_land.csv", sep = ""), 
                 head = TRUE, sep = ",", stringsAsFactors = FALSE) %>% 
  mutate(CATCH = CATCH * 1000)

Length_Weight <- read.csv(paste(path, "/final_output/nafo_LengthWeight.csv", sep = ""),
                          header = TRUE, sep = ",",
                          stringsAsFactors = FALSE) # read in length-at-weight data for ESS and WSS


#  Most indicators --------------------------------------------------------

inds <- extractAll(X = RV, 
                   X_length = RV_length, 
                   land = land, 
                   years = ind.years,
                   speciesinfo.table = species.info,
                   species.table = species.groups,
                   metric.bio = "ABUNDANCE", 
                   group.bio = c("ALL", "GROUNDFISH"), 
                   minTL.bio = 3,
                   LSI.group = "ALL", 
                   max.length = 85, 
                  LFI.group = "ALL", 
                   large.fish = 35,
                            LenWt.table = Length_Weight,
                            condition.groups = condition.groups, 
                            guild.groups = trophicguild.groups,
                            ratio.groups = ratio.groups,
                            maxlength.group = "FINFISH",  
                            TL.grouping = 1, 
                            wind = 5, 
                            negative = FALSE,
                            resource.groups = resource.groups,
                            minTL.FiB = 0, 
                            base.start = 1968, 
                            base.end = 1970, 
                            TE = 0.1,
                            landings.groups = landings.groups, 
                            FP.groups = FP.groups, 
                            minTL.FP = c(0, 3.25),
                            glob.env = TRUE,
                            raw = TRUE, 
                            std = FALSE, export.path = NULL, export.id = area)

# remove 2018 from ESS data because survey was not completed that year
# inds.ess <- inds %>% filter(ID != "4X", YEAR != 2018)
# inds.wss <- inds %>% filter(ID == "4X")
# inds <- rbind(inds.ess, inds.wss)


#the extractAll produces error above. 
#Biodiversity, Structure and Functioning, Stability and Resistance, Resource Potential, and Fishing Pressure

# allBiodiversity
inds_biodiv<-allBiodiversity(X=RV,
                             metric="ABUNDANCE", 
                             groups="ALL",
                             percentiles= c(0.25, 0.75),
                             minTL=0, 
                             TL.table=NULL,
                             year=ind.years,
                             raw=TRUE,
                             std=TRUE,
                             glob.env=TRUE
                             )


inds_struct<-allStructure( X=RV,
  X_length=RV_length,
  LSI.group="ALL",
  LFI.group="ALL",
  max.length = 85,
  large.fish = 35,
  guild.groups=trophicguild.groups,
  condition.groups=condition.groups,
  ratio.groups=ratio.groups,
  species.table=species.groups,
  speciesinfo.table=species.table,
  LenWt.table=Length_Weight,
  years=ind.years,
  raw = TRUE,
  std = TRUE,
  glob.env = TRUE,
  export.path = NULL,
  export.id = NULL
)

inds_stab<-allStability(X=RV,
  land=land,
  maxlength.group = "FINFISH",
  species.table = species.groups,
  speciesinfo.table = species.info,
  TL.grouping = 1,
  wind = 5,
  negative = FALSE,
  years=ind.years,
  raw = TRUE,
  std = TRUE,
  glob.env = TRUE,
  export.path = NULL,
  export.id = NULL
)

inds_press<-allPressure( X=RV,
  land=land,
  species.table = species.groups,
  speciesinfo.table = species.info,
  landings.groups = landings.groups,
  FP.groups = FP.groups,
  minTL = c(0, 3.25),
  years=ind.years,
  raw = TRUE,
  std = TRUE,
  glob.env = TRUE,
  export.path = NULL,
  export.id = NULL
)

inds_pote<-allPotential(X=RV,
  land = land,
  species.table = species.groups,
  speciesinfo.table= species.info,
  resource.groups = resource.groups,
  minTL = 0,
  TE = 0.1,
  base.start= 1968,
  base.end = 1970,
  years = ind.years,
  raw = TRUE,
  std = TRUE,
  glob.env = TRUE,
  export.path = NULL,
  export.id = NULL
  )


# save YEAR and ID to add into inds_imp dataframe
YEAR <- inds$YEAR
ID <- inds$ID

# BIOMASS_TL2 because only using values after 1998 (added back in to allIndicators below)
BIOMASS_TL2 <- inds %>% 
  dplyr::select(YEAR, ID, BIOMASS_TL2) %>% 
  filter(YEAR >= 1999)

# remove BIOMASS_TL2 from inds so it is not used in the imputation (consistent with esswss)
inds <- inds %>% dplyr::select(-BIOMASS_TL2)

# record which indicators have >25% NA for given area
# these indicators will be filtered out after the imputation
NA_index <- get_NA_index(inds)
NA_index_output <- pivot_wider(NA_index, names_from = col_index, values_from = fraction_NA)
names(NA_index_output)[1] = "NAFO_Area"
write.csv(NA_index_output, file = paste(export.filepath, "/fraction_NA_", area, ".csv", sep = ""),
          row.names = FALSE)

# imputation to fill in missing values
set.seed(22)
inds_imp <- inds %>% 
  dplyr::select(-YEAR, -ID) %>% 
  as.matrix() %>% 
  missForest()

inds_imp <- inds_imp$ximp %>% 
  data.frame() %>% 
  mutate(YEAR = YEAR, ID = ID) %>% 
  dplyr::select(YEAR, ID, everything())

# filter out indicators with >25% NA for a given area
# this function uses the NA_index dataframe to find indicators for each strata that
## had >25% NAs before imputation. All values for these indicators are replaced with
## NA, essentially filtering out these indicators
inds_filtered <- add_NAs(inds_imp, NA_index)
rm(inds, inds_imp)


# Invertebrate indicators -------------------------------------------------

# the above calculates everything EXCEPT invertebrate-related indicators 
## (because we want to start the invertebrate indicators in 1999)

# invertebrate biomass
inverts <- resourcePotential(RV, groups = "INVERTEBRATES", species.table = species.groups, 
                             metric = "BIOMASS", years = c(1999:end_year))

# invertebrate to demersal ratio
invert_dem <- data.frame(rbind(c("INVERTEBRATES", "GROUNDFISH")))
names(invert_dem) <- c("group1", "group2")
invert_demersal_ratio <- biomassRatio(RV, ratio.groups = invert_dem, 
                                      species.table = species.groups, 
                                      metric = "BIOMASS", years = c(1999:end_year))

# fishing pressure on inverts
FP.groups.inverts <- data.frame(rbind(c("INVERTEBRATES", "INVERTEBRATES.L")))
names(FP.groups.inverts) <- c("group.X", "group.land")
FP_inverts <- fishingPressure(RV, land, FP.groups = FP.groups.inverts, species.table = species.groups,
                              years = c(1999:end_year))

short_inds <- full_join(inverts, invert_demersal_ratio, by = c("YEAR", "ID")) %>% 
  full_join(FP_inverts, by = c("YEAR", "ID")) %>% 
  full_join(BIOMASS_TL2, by = c("YEAR", "ID"))

# remove 2018 from ESS data because survey was not completed that year
# short_inds.ess <- short_inds %>% filter(ID != "4X", YEAR != 2018)
# short_inds.wss <- short_inds %>% filter(ID == "4X")
# short_inds <- rbind(short_inds.ess, short_inds.wss)
# rm(short_inds.ess, short_inds.wss)

# record which indicators have >25% NA for given strata
# these indicators will be filtered out after the imputation
NA_index_short <- get_NA_index(short_inds)

# Export -----------------------------------------------------------------

allIndicators <- full_join(inds_filtered, short_inds, by = c("YEAR", "ID")) 

write.csv(allIndicators, file = paste(export.filepath, "/eco_indicators_", area, ".csv", sep = ""),
          row.names = FALSE)


```
                          
                          