# setup
```{r}

library(tidyverse)
library(mgcv)
library(gratia)
library(ggplot2)
library(patchwork)
library(marea)
# library(gslea)
# library(azmpdata)
```

#load saved data
```{r}
load("D:/Work projects/MAReco_indicators/data_thresholds/ess_data_longts.Rdata")

load("D:/Work projects/MAReco_indicators/data_thresholds/wss_data_longts.Rdata")
```


# set region and model selection rules 
```{r}

# --- Global model options (consistent across GLM/GAM/tGAM) ---
fit_method <- "REML"
use_select <- TRUE
gamma_val  <- 1.4

# --- Selection rule parameters ---
tolerance <- 0.01  # 1% REML tolerance
alpha     <- 0.08  # hinge significance threshold (TGAM)

# --- Bootstrap controls (serial; no parallel) ---
set.seed(20240229)
n_boot       <- 100   # increase to 500+ for final results
min_success  <- 50
n_grid       <- 300
min_run      <- 5
trim_bd      <- 0.05  # for max-curvature edge trim


# ---- Helper: robust column accessor (gratia version-proof) ----*maybe move to .R*
get_col <- function(df, candidates) {
  for (nm in candidates) if (nm %in% names(df)) return(df[[nm]])
  stop("None of the candidate columns found: ", paste(candidates, collapse = ", "))
}

# ---- Helper: prediction grid builder ----
build_grid <- function(model, pred, n = 400) {
  stopifnot(pred %in% names(model$model))
  rng <- range(model$model[[pred]], na.rm = TRUE)
  setNames(data.frame(seq(rng[1], rng[2], length.out = n)), pred)
}


# ---- Effective thresholds on a GAM ----
effective_derivative_onset <- function(model, predictor, min_run = 5, n = 400) {
  nd <- build_grid(model, predictor, n)
  der <- try(
    derivatives(model, term = paste0("s(", predictor, ")"),
                order = 1, type = "link", interval = "confidence",
                newdata = nd, partial_match = TRUE),
    silent = TRUE
  )
  if (inherits(der, "try-error")) return(tibble(Deriv_Onset = NA_real_, Deriv_Onset_Sign = NA_character_))
  x     <- get_col(der, c(predictor))
  lower <- get_col(der, c("lower", "lower_ci"))
  upper <- get_col(der, c("upper", "upper_ci"))
  sig <- ifelse(lower > 0,  1L, ifelse(upper < 0, -1L, 0L))
  rl <- rle(sig)
  pos <- 0L; onset_x <- NA_real_; onset_sign <- NA_character_
  for (i in seq_along(rl$lengths)) {
    len <- rl$lengths[i]; val <- rl$values[i]
    idx_start <- pos + 1L; idx_end <- pos + len; pos <- idx_end
    if (val != 0L && len >= min_run) {
      onset_x <- x[idx_start]
      onset_sign <- if (val == 1L) "positive" else "negative"
      break
    }
  }
  tibble(Deriv_Onset = onset_x, Deriv_Onset_Sign = onset_sign)
}

effective_max_curvature <- function(model, predictor, trim = 0.05, n = 400) {
  nd <- build_grid(model, predictor, n)
  der2 <- try(
    derivatives(model, term = paste0("s(", predictor, ")"),
                order = 2, type = "link", interval = "none",
                newdata = nd, partial_match = TRUE),
    silent = TRUE
  )
  if (inherits(der2, "try-error")) return(tibble(Max_Curvature_x = NA_real_, Max_Curvature_val = NA_real_))
  x  <- get_col(der2, c(predictor))
  d2 <- get_col(der2, c(".derivative", "derivative"))
  k <- length(x)
  i1 <- max(1, floor(trim * k))
  i2 <- min(k,  ceiling((1 - trim) * k))
  if (i2 <= i1) return(tibble(Max_Curvature_x = NA_real_, Max_Curvature_val = NA_real_))
  sl <- seq(i1, i2)
  idx <- sl[which.max(abs(d2[sl]))]
  tibble(Max_Curvature_x = x[idx], Max_Curvature_val = d2[idx])
}

compute_effective_thresholds <- function(model, predictor, min_run = 5, trim = 0.05, n = 400) {
  a <- effective_derivative_onset(model, predictor, min_run = min_run, n = n)
  b <- effective_max_curvature(model, predictor, trim = trim, n = n)
  bind_cols(a, b)
}

# ---- Robust derivative zero-crossings (GAM) using CI ----
robust_gam_crossings <- function(model, predictor, n = 300) {
  if (!predictor %in% names(model$model)) return(NULL)
  x_min <- min(model$model[[predictor]], na.rm = TRUE)
  x_max <- max(model$model[[predictor]], na.rm = TRUE)
  if (!is.finite(x_min) || !is.finite(x_max) || x_min == x_max) return(NULL)
  nd <- setNames(data.frame(seq(x_min, x_max, length.out = n)), predictor)
  deriv <- try(
    derivatives(model, term = paste0("s(", predictor, ")"),
                order = 1, type = "link", interval = "confidence",
                newdata = nd, partial_match = TRUE),
    silent = TRUE
  )
  if (inherits(deriv, "try-error")) return(NULL)
  x     <- get_col(deriv, c(predictor))
  d     <- get_col(deriv, c(".derivative", "derivative"))
  lower <- get_col(deriv, c("lower", "lower_ci"))
  upper <- get_col(deriv, c("upper", "upper_ci"))
  sig_class <- ifelse(lower > 0, 1L, ifelse(upper < 0, -1L, 0L))
  idx <- which(diff(sig_class) != 0 & sig_class[-length(sig_class)] != 0 & sig_class[-1] != 0)
  if (length(idx) == 0) return(NULL)
  crossings <- purrr::map_dbl(idx, function(i) {
    x1 <- x[i]; x2 <- x[i + 1]; d1 <- d[i]; d2 <- d[i + 1]
    if (!is.finite(d2 - d1) || (d2 - d1) == 0) return(NA_real_)
    x1 - d1 * (x2 - x1) / (d2 - d1)
  })
  crossings <- crossings[is.finite(crossings)]
  if (length(crossings) == 0) return(NULL)
  tibble(Crossing_Index = seq_along(crossings),
         Derivative_Zero_Crossing = crossings)
}

# ---- tGAM: grid-search hinge (continuous kink) ----
fit_threshold_gam <- function(df, resp, pres,
                              family = tw(),
                              method = fit_method,
                              select = use_select,
                              gamma  = gamma_val,
                              k_main = NULL, k_hinge = NULL,
                              probs_grid = seq(0.15, 0.85, by = 0.02),
                              min_per_side = 6) {
  n <- nrow(df)
  if (is.null(k_main))  k_main  <- max(5, min(10, floor(n / 4)))
  if (is.null(k_hinge)) k_hinge <- max(5, min(10, floor(n / 4)))
  x <- df[[pres]]
  cands <- unique(quantile(x, probs = probs_grid, na.rm = TRUE, names = FALSE))
  cands <- cands[is.finite(cands)]
  cands <- cands[vapply(cands, function(c0) {
    sum(x <= c0, na.rm = TRUE) >= min_per_side && sum(x > c0, na.rm = TRUE) >= min_per_side
  }, logical(1))]
  if (length(cands) == 0L) return(NULL)

  best <- NULL; best_score <- Inf; best_c <- NA_real_
  for (c0 in cands) {
    df$hinge <- pmax(0, df[[pres]] - c0)
    fml <- as.formula(paste0(resp, " ~ s(", pres, ", k=", k_main, ") + s(hinge, k=", k_hinge, ")"))
    fit <- try(
      gam(fml, data = df, family = family, method = method, select = select, gamma = gamma),
      silent = TRUE
    )
    if (inherits(fit, "try-error")) next
    score <- AIC(fit)
    if (!is.finite(score)) next
    if (score < best_score) { best <- fit; best_score <- score; best_c <- c0 }
  }
  if (is.null(best)) return(NULL)
  attr(best, "threshold") <- best_c
  best
}

# ---- Bootstrap (slope) global controls ----
use_block_boot <- TRUE    # TRUE recommended for time series
block_size     <- 5       # choose based on ACF/seasonality
B_slope        <- 300     # 300 for development; 500–1000 for final

# ---- Helpers: Paired moving-block bootstrap indices ----
paired_mbb_indices <- function(n, B, block_size = 5) {
  if (is.null(block_size) || block_size <= 1L) {
    replicate(B, sample.int(n, replace = TRUE), simplify = FALSE)
  } else {
    sample_blocks <- function() {
      starts <- sample.int(n - block_size + 1L, size = ceiling(n / block_size), replace = TRUE)
      idx <- unlist(Map(seq.int, from = starts, to = starts + block_size - 1L))
      idx[seq_len(n)]
    }
    replicate(B, sample_blocks(), simplify = FALSE)
  }
}

# ---- Helper: get derivative column robustly (gratia version-proof) ----
get_deriv_col <- function(df) {
  if (".derivative" %in% names(df)) return(df[[".derivative"]])
  if ("derivative"  %in% names(df)) return(df[["derivative"]])
  stop("Could not find derivative column from gratia::derivatives() output.")
}

# ---- Bootstrap the GAM slope (first derivative) on a common grid ----
bootstrap_gam_slope <- function(model, predictor, B = 300, block_size = 5, n_grid = 300) {
  stopifnot(inherits(model, "gam"))
  df <- model$model
  n  <- nrow(df)
  if (!predictor %in% names(df) || n < 10) return(NULL)

  x_min <- min(df[[predictor]], na.rm = TRUE)
  x_max <- max(df[[predictor]], na.rm = TRUE)
  nd    <- setNames(data.frame(seq(x_min, x_max, length.out = n_grid)), predictor)

  idx_list <- paired_mbb_indices(n, B, if (use_block_boot) block_size else 1L)
  d1_mat <- matrix(NA_real_, nrow = n_grid, ncol = B)

  for (b in seq_len(B)) {
    df_b <- df[idx_list[[b]], , drop = FALSE]
    fit_b <- try(
      mgcv::gam(formula(model), data = df_b,
                family = model$family, method = model$method,
                select = TRUE, gamma = model$gamma),
      silent = TRUE
    )
    if (inherits(fit_b, "try-error")) next

    d1_b <- try(
      gratia::derivatives(
        fit_b, term = paste0("s(", predictor, ")"),
        order = 1, type = "link", interval = "none",
        newdata = nd, partial_match = TRUE
      ),
      silent = TRUE
    )
    if (!inherits(d1_b, "try-error")) {
      d1_mat[, b] <- get_deriv_col(d1_b)
    }
  }

  # Quantiles per grid point (require at least 20% successful draws)
  qfun <- function(v) {
    apply(v, 1, function(z) {
      z <- z[is.finite(z)]
      if (length(z) < ceiling(0.2 * B)) return(c(NA, NA, NA))
      stats::quantile(z, c(0.025, 0.5, 0.975), na.rm = TRUE)
    })
  }
  q1 <- qfun(d1_mat)
  tibble::tibble(
    !!predictor := nd[[predictor]],
    slope_q025 = q1[1, ], slope_med = q1[2, ], slope_q975 = q1[3, ]
  )
}

# ---- Helper: slope onset detector (first run of significant slope) ----
# Returns first x where the slope CI excludes 0 for >= min_run consecutive grid points
slope_onset_from_ci <- function(slope_df, xvar, min_run = 5) {
  stopifnot(all(c("slope_q025", "slope_q975") %in% names(slope_df)))
  x   <- slope_df[[xvar]]
  low <- slope_df$slope_q025
  up  <- slope_df$slope_q975
  sig <- ifelse(low > 0,  1L, ifelse(up < 0, -1L, 0L))
  rl  <- rle(sig)
  pos <- 0L
  for (i in seq_along(rl$lengths)) {
    len <- rl$lengths[i]; val <- rl$values[i]
    idx_start <- pos + 1L; pos <- pos + len
    if (val != 0L && len >= min_run) {
      return(tibble::tibble(
        Slope_Onset_x = x[idx_start],
        Slope_Onset_Sign = if (val == 1L) "positive" else "negative"
      ))
    }
  }
  tibble::tibble(Slope_Onset_x = NA_real_, Slope_Onset_Sign = NA_character_)
}

# ---- Helper: zero-crossings of the bootstrap-median slope (linear interpolation) ----
slope_zero_crossings <- function(slope_df, xvar) {
  x <- slope_df[[xvar]]
  y <- slope_df$slope_med
  if (length(x) < 2 || sum(is.finite(y)) < 2) return(numeric(0))
  idx <- which(diff(sign(y)) != 0 & is.finite(y[-length(y)]) & is.finite(y[-1]))
  if (!length(idx)) return(numeric(0))
  crossings <- purrr::map_dbl(idx, function(i) {
    x1 <- x[i]; x2 <- x[i + 1]; y1 <- y[i]; y2 <- y[i + 1]
    if (!is.finite(y2 - y1) || (y2 - y1) == 0) return(NA_real_)
    x1 - y1 * (x2 - x1) / (y2 - y1)
  })
  crossings[is.finite(crossings)]
}
```

# indicator pressure setup
```{r}
# --- Region label (used in filenames) ---
region <- "wss"

indicators <- wss_data_longts$indicators |> 
  select(-year)
pressures  <- wss_data_longts$pressures |> select(-year)
```

# Step 1. Fit GLM GAM tGAM with grid search for correlations 
```{r}
# ---- Step 1: Fit GLM, GAM, and tGAM models ----

model_list <- list()

for (a_col in names(indicators)) {

  for (b_col in names(pressures)) {

    tryCatch({

      # Create paired dataset
      df <- data.frame(
        resp = indicators[[a_col]],
        pres = pressures[[b_col]]
      )

      names(df) <- c(a_col, b_col)

      df <- na.omit(df)

      # Basic data checks
      if (nrow(df) < 10) next

      if (var(df[[a_col]], na.rm = TRUE) == 0) next
      if (var(df[[b_col]], na.rm = TRUE) == 0) next

      # -------------------------------
      # GAM
      # -------------------------------
      gam_model <- mgcv::gam(
        as.formula(
          paste0(a_col, " ~ s(", b_col, ")")
        ),
        data = df,
        family = tw(),
        method = fit_method,
        select = use_select,
        gamma = gamma_val
      )

      # -------------------------------
      # GLM (fit through gam)
      # -------------------------------
      glm_model <- mgcv::gam(
        as.formula(
          paste0(a_col, " ~ ", b_col)
        ),
        data = df,
        family = tw(),
        method = fit_method,
        select = use_select,
        gamma = gamma_val
      )

      # -------------------------------
      # Threshold GAM
      # -------------------------------
      tgam_model <- fit_threshold_gam(
        df,
        resp   = a_col,
        pres   = b_col,
        family = tw()
      )

      # -------------------------------
      # Save models
      # -------------------------------
      model_base <- paste(a_col, b_col, sep = "_vs_")

      model_list[[paste0(model_base, "_GAM")]] <- gam_model
      model_list[[paste0(model_base, "_GLM")]] <- glm_model

      if (!is.null(tgam_model)) {
        model_list[[paste0(model_base, "_TGAM")]] <- tgam_model
      }

    }, error = function(e) {

      message(
        "Error fitting models for ",
        a_col, " vs ", b_col,
        ": ", conditionMessage(e)
      )

    })
  }
}

message("Finished fitting ", length(model_list), " models.")


```


#Step 2. Extract model diagonstics 
```{r}
# ---- Step 2: Diagnostics ----
comparison_results <- data.frame()

for (model_name in names(model_list)) {
  tryCatch({
    model <- model_list[[model_name]]
    smry  <- summary(model)

    parts <- strsplit(model_name, "_vs_")[[1]]
    response <- parts[1]
    predictor_and_type <- parts[2]
    predictor <- sub("_(GAM|GLM|TGAM)$", "", predictor_and_type)
    model_type <- sub("^.*_(GAM|GLM|TGAM)$", "\\1", predictor_and_type)

    aic_val   <- AIC(model)
    r_squared <- smry$r.sq
    gcv_ubre  <- model$gcv.ubre
    edf_total <- tryCatch({
      if (!is.null(smry$s.table)) sum(smry$s.table[, "edf"], na.rm = TRUE)
      else if (!is.null(model$edf)) sum(model$edf, na.rm = TRUE)
      else NA_real_
    }, error = function(e) NA_real_)

    p_value <- tryCatch({
      if (model_type == "GAM") {
        smry$s.table[1, "p-value"]
      } else if (model_type == "GLM") {
        if (predictor %in% rownames(smry$p.table)) smry$p.table[predictor, "Pr(>|t|)"]
        else smry$p.table[2, "Pr(>|t|)"]
      } else if (model_type == "TGAM") {
        if (!is.null(smry$s.table) && "s(hinge)" %in% rownames(smry$s.table)) {
          smry$s.table["s(hinge)", "p-value"]
        } else if (!is.null(smry$s.table)) {
          min(smry$s.table[, "p-value"], na.rm = TRUE)
        } else NA
      } else NA
    }, error = function(e) NA)

    threshold <- if (model_type == "TGAM") attr(model, "threshold") else NA_real_

    comparison_results <- rbind(
      comparison_results,
      data.frame(
        Response   = response,
        Predictor  = predictor,
        Model_Type = model_type,
        AIC        = aic_val,
        R_squared  = r_squared,
        EDF        = edf_total,
        GCV_UBRE   = gcv_ubre,
        p_value    = p_value,
        Threshold  = threshold,
        stringsAsFactors = FALSE
      )
    )
  }, error = function(e) {
    message("Error extracting diagnostics for ", model_name, ": ", e$message)
  })
}

print(comparison_results)
```

# Step 3. Model Selection
```{r}
# ---- Step 3: Strict single-model selection ----

model_selection <- comparison_results %>%
  filter(Model_Type %in% c("GLM", "GAM", "TGAM")) %>%
  group_by(Response, Predictor) %>%
  mutate(
    # Identify best GAM/GLM
    best_non_tgam_aic = min(AIC[Model_Type != "TGAM"], na.rm = TRUE),

    # Identify best TGAM
    tgam_aic = ifelse(Model_Type == "TGAM", AIC, NA_real_),
    tgam_p   = ifelse(Model_Type == "TGAM", p_value, NA_real_),

    tgam_aic_best = min(tgam_aic, na.rm = TRUE),
    tgam_p_best   = min(tgam_p, na.rm = TRUE),

    # Compute delta AIC
    delta_aic = tgam_aic_best - best_non_tgam_aic,

    # Decision: only accept TGAM if justified
    pick_tgam = !is.infinite(tgam_aic_best) & (
      delta_aic <= -2 |
      (delta_aic <= 2 & tgam_p_best < alpha)
    )
  ) %>%

  summarise(
    # ✅ pick ONE model per pair
    Model_Type = if (any(pick_tgam, na.rm = TRUE)) {
      "TGAM"
    } else {
      Model_Type[which.min(AIC[Model_Type != "TGAM"])]
    },

    # ✅ grab row corresponding to selected model
    AIC        = AIC[Model_Type == cur_data()$Model_Type][1],
    R_squared  = R_squared[Model_Type == cur_data()$Model_Type][1],
    EDF        = EDF[Model_Type == cur_data()$Model_Type][1],
    GCV_UBRE   = GCV_UBRE[Model_Type == cur_data()$Model_Type][1],
    p_value    = p_value[Model_Type == cur_data()$Model_Type][1],

    .groups = "drop"
  )


print(model_selection)
```

# Step 4. Threshold analysis

```{r}
# ---- Step 4: Robust GAM derivative zero-crossings + tGAM threshold extraction (patched) ----
library(gratia)
library(dplyr)
library(purrr)
library(tidyr)

# Helper: get a column by candidate names (for cross-version compatibility)
get_col <- function(df, candidates) {
  for (nm in candidates) {
    if (nm %in% names(df)) return(df[[nm]])
  }
  stop("None of the candidate columns found: ", paste(candidates, collapse = ", "))
}

# Robust zero-crossing function for GAMs:
# - builds newdata grid so predictor column is present
# - uses CI on derivative
# - finds transitions only between significant + and -
robust_gam_crossings <- function(model, predictor, n = 300) {
  # Guard: predictor must be in model frame
  if (!predictor %in% names(model$model)) return(NULL)
  
  x_min <- min(model$model[[predictor]], na.rm = TRUE)
  x_max <- max(model$model[[predictor]], na.rm = TRUE)
  if (!is.finite(x_min) || !is.finite(x_max) || x_min == x_max) return(NULL)
  
  nd <- setNames(data.frame(seq(x_min, x_max, length.out = n)), predictor)
  
  # Compute first derivative WITH CI on the link scale and guaranteed predictor column
  deriv <- try(
    derivatives(
      model,
      term = paste0("s(", predictor, ")"),
      order = 1,
      type = "link",
      interval = "confidence",
      n = n,
      newdata = nd,
      partial_match = TRUE
    ),
    silent = TRUE
  )
  if (inherits(deriv, "try-error")) return(NULL)
  
  # Extract columns robustly
  x     <- get_col(deriv, c(predictor))                      # predictor column guaranteed by newdata
  d     <- get_col(deriv, c(".derivative", "derivative"))    # derivative column name varies by gratia
  lower <- get_col(deriv, c("lower", "lower_ci"))
  upper <- get_col(deriv, c("upper", "upper_ci"))
  
  # Classify derivative significance
  sig_class <- ifelse(lower > 0,  1L,
               ifelse(upper < 0, -1L, 0L))
  
  # Look for transitions between significant regions only
  idx <- which(diff(sig_class) != 0 &
               sig_class[-length(sig_class)] != 0 &
               sig_class[-1] != 0)
  if (length(idx) == 0) return(NULL)
  
  # Linear interpolation of the zero of d between x[i] and x[i+1]
  crossings <- map_dbl(idx, function(i) {
    x1 <- x[i];   x2 <- x[i + 1]
    d1 <- d[i];   d2 <- d[i + 1]
    if (!is.finite(d2 - d1) || (d2 - d1) == 0) return(NA_real_)
    x1 - d1 * (x2 - x1) / (d2 - d1)
  })
  crossings <- crossings[is.finite(crossings)]
  if (length(crossings) == 0) return(NULL)
  
  tibble(
    Crossing_Index = seq_along(crossings),
    Derivative_Zero_Crossing = crossings
  )
}

# Containers
gam_crossings_list <- list()
tgam_threshold_list <- list()

# Iterate through models
for (model_name in names(model_list)) {
  model <- model_list[[model_name]]
  # Defensive: skip if not a gam object
  if (!inherits(model, "gam")) next
  
  # Parse names
  parts <- strsplit(model_name, "_vs_")[[1]]
  response <- parts[1]
  predictor_and_type <- parts[2]
  predictor <- sub("_(GAM|GLM|TGAM)$", "", predictor_and_type)
  model_type <- sub("^.*_(GAM|GLM|TGAM)$", "\\1", predictor_and_type)
  
  # --- A) TGAM: extract threshold
  if (identical(model_type, "TGAM")) {
    thr <- attr(model, "threshold")
    if (!is.null(thr) && is.finite(thr)) {
      tgam_threshold_list[[model_name]] <- tibble(
        Model = model_name,
        Response = response,
        Predictor = predictor,
        tGAM_Threshold = thr
      )
    }
  }
  
  # --- B) GAM: robust derivative zero-crossings
  if (identical(model_type, "GAM")) {
    cr <- robust_gam_crossings(model, predictor, n = 300)
    if (!is.null(cr)) {
      gam_crossings_list[[model_name]] <- cr %>%
        mutate(
          Model = model_name,
          Response = response,
          Predictor = predictor
        )
    }
  }
}

# Bind with safety (empty -> 0-row tibble with expected cols)
crossings_df <- if (length(gam_crossings_list)) {
  bind_rows(gam_crossings_list) %>%
    select(Model, Response, Predictor, Crossing_Index, Derivative_Zero_Crossing) %>%
    arrange(Response, Predictor, Crossing_Index)
} else {
  tibble(Model = character(), Response = character(), Predictor = character(),
         Crossing_Index = integer(), Derivative_Zero_Crossing = numeric())
}

threshold_df <- if (length(tgam_threshold_list)) {
  bind_rows(tgam_threshold_list) %>%
    arrange(Response, Predictor)
} else {
  tibble(Model = character(), Response = character(), Predictor = character(),
         tGAM_Threshold = numeric())
}

# Unified table (one row per model; crossings collapsed)
thresholds_all_df <- full_join(
  crossings_df %>%
    group_by(Model, Response, Predictor) %>%
    summarise(Derivative_Zero_Crossing = paste(signif(Derivative_Zero_Crossing, 4), collapse = ", "),
              .groups = "drop"),
  threshold_df,
  by = c("Model", "Response", "Predictor")
) %>%
  arrange(Response, Predictor)

print(thresholds_all_df)

# Fill missing entries with NA (for models that are GAM-only or TGAM-only)
thresholds_all_df <- thresholds_all_df %>%
  arrange(Response, Predictor)

print(thresholds_all_df)


```

# Step 5. Compute slope CIs for selected models
```{r}
# ---- Step 5: Bootstrap slopes for selected GAMs ----
selected_only <- TRUE #select FALSE for all models

target_gams <- if (selected_only) {
  model_selection %>% filter(Model_Type == "GAM") %>%
    transmute(Key = paste0(Response, "_vs_", Predictor, "_", Model_Type))
} else {
  tibble::tibble(Key = names(model_list)[grepl("_GAM$", names(model_list))])
}

gam_slope_boot_list <- list()
gam_slope_boot_summary_list <- list()

for (key in target_gams$Key) {
  model <- model_list[[key]]
  if (is.null(model) || !inherits(model, "gam")) next

  parts <- strsplit(key, "_vs_")[[1]]
  response <- parts[1]
  predictor <- sub("_.*$", "", parts[2])

  slope_df <- try(
    bootstrap_gam_slope(model, predictor,
                        B = B_slope, block_size = block_size, n_grid = n_grid),
    silent = TRUE
  )
  if (inherits(slope_df, "try-error") || is.null(slope_df)) next

  slope_df$Model     <- key
  slope_df$Response  <- response
  slope_df$Predictor <- predictor

  gam_slope_boot_list[[key]] <- slope_df

  # Extract onset and crossings
  onset <- slope_onset_from_ci(slope_df, predictor, min_run = min_run)
  crossings <- slope_zero_crossings(slope_df, predictor)
  gam_slope_boot_summary_list[[key]] <- tibble::tibble(
    Model = key, Response = response, Predictor = predictor,
    Slope_Onset_x = onset$Slope_Onset_x,
    Slope_Onset_Sign = onset$Slope_Onset_Sign,
    Slope_Zero_Crossings = if (length(crossings)) paste(signif(crossings, 4), collapse = ", ") else NA_character_
  )
}

gam_slope_boot_df <- if (length(gam_slope_boot_list)) dplyr::bind_rows(gam_slope_boot_list) else
  tibble::tibble(Model=character(), Response=character(), Predictor=character(),
                 slope_q025=numeric(), slope_med=numeric(), slope_q975=numeric())

gam_slope_boot_summary_df <- if (length(gam_slope_boot_summary_list)) dplyr::bind_rows(gam_slope_boot_summary_list) else
  tibble::tibble(Model=character(), Response=character(), Predictor=character(),
                 Slope_Onset_x=numeric(), Slope_Onset_Sign=character(), Slope_Zero_Crossings=character())

print(gam_slope_boot_summary_df)
```


#Step 6.Build best model tables
```{r}
best_models_table <- model_selection %>%
  rowwise() %>%
  mutate(
    Model_Key = paste0(Response, "_vs_", Predictor, "_", Model_Type),

    Threshold = if (Model_Type == "TGAM") {
      if (!Model_Key %in% names(model_list)) {
        NA_real_
      } else {
        thr <- attr(model_list[[Model_Key]], "threshold")
        if (is.null(thr) || !is.finite(thr)) NA_real_ else as.numeric(thr)
      }
    } else NA_real_,

    Derivative_Zero_Crossings = if (Model_Type == "GAM") {
      if (exists("crossings_df") && nrow(crossings_df) > 0) {
        collapse_crossings(Model_Key, crossings_df)
      } else {
        NA_character_
      }
    } else NA_character_
  ) %>%
  ungroup() %>%
  select(Response, Predictor, Model_Type,
         AIC, R_squared, EDF, GCV_UBRE, p_value,
         Threshold, Derivative_Zero_Crossings) %>%
  arrange(Response, Predictor)

print(best_models_table)
```

#Step 7. Unified thresholds table
```{r}
# ---- Step 5b: Unified threshold table (tGAM + GAM slope + GAM effective) ----

# Ensure required objects exist (safe empties)
if (!exists("tgam_boot_summary_df")) tgam_boot_summary_df <- tibble::tibble(
  Model=character(), Response=character(), Predictor=character(),
  c_q025=numeric(), c_med=numeric(), c_q975=numeric(), n_success=integer()
)
if (!exists("effective_thresholds_df")) effective_thresholds_df <- tibble::tibble(
  Model=character(), Response=character(), Predictor=character(),
  Deriv_Onset=numeric(), Deriv_Onset_Sign=character(),
  Max_Curvature_x=numeric(), Max_Curvature_val=numeric()
)
if (!exists("crossings_df")) crossings_df <- tibble::tibble(
  Model=character(), Response=character(), Predictor=character(),
  Crossing_Index=integer(), Derivative_Zero_Crossing=numeric()
)
if (!exists("gam_slope_boot_summary_df")) gam_slope_boot_summary_df <- tibble::tibble(
  Model=character(), Response=character(), Predictor=character(),
  Slope_Onset_x=numeric(), Slope_Onset_Sign=character(), Slope_Zero_Crossings=character()
)

unified_thresholds <- model_selection %>%
  mutate(Model_Key = paste0(Response, "_vs_", Predictor, "_", Model_Type)) %>%
  left_join(
    tgam_boot_summary_df %>%
      transmute(Model, tGAM_Threshold_Med = c_med, tGAM_Threshold_q025 = c_q025, tGAM_Threshold_q975 = c_q975),
    by = c("Model_Key" = "Model")
  ) %>%
  left_join(
    gam_slope_boot_summary_df %>%
      transmute(Model, Slope_Onset_x, Slope_Onset_Sign, Slope_Zero_Crossings),
    by = c("Model_Key" = "Model")
  ) %>%
  left_join(
    effective_thresholds_df %>%
      transmute(Model, Deriv_Onset, Deriv_Onset_Sign, Max_Curvature_x),
    by = c("Model_Key" = "Model")
  ) %>%
  left_join(
    crossings_df %>%
      group_by(Model) %>%
      summarise(Deriv_Zero_Crossings = paste(signif(Derivative_Zero_Crossing, 4), collapse = ", "),
                .groups = "drop"),
    by = c("Model_Key" = "Model")
  ) %>%
  select(Response, Predictor, Model_Type,
         # tGAM hinge
         tGAM_Threshold_Med, tGAM_Threshold_q025, tGAM_Threshold_q975,
         # GAM slope bootstrap
         Slope_Onset_x, Slope_Onset_Sign, Slope_Zero_Crossings,
         # GAM analytical thresholds
         Deriv_Onset, Deriv_Onset_Sign, Max_Curvature_x,
         Deriv_Zero_Crossings,
         # Diagnostics (from selection frame)
         AIC, R_squared, EDF, GCV_UBRE, p_value) %>%
  arrange(Response, Predictor)

print(unified_thresholds)
```
# Step 8. Plotting
```{r}
# ---- Step 8: Plot ONLY selected best models (organized by pressure) ----
library(ggplot2)
library(patchwork)
library(dplyr)
library(here)

plot_list <- list()

# ✅ Build plotting table and sort by pressure first
# plot_df <- model_selection %>%
#   mutate(
#     Model_Key = paste0(Response, "_vs_", Predictor, "_", Model_Type)
#   ) %>%
#   arrange(Predictor, Response)

plot_df <- model_selection %>%
  filter(!is.na(p_value), p_value < alpha) %>%   # ✅ KEEP ONLY SIGNIFICANT
  mutate(
    Model_Key = paste0(Response, "_vs_", Predictor, "_", Model_Type)
  ) %>%
  arrange(Predictor, Response)


selected_keys <- unique(plot_df$Model_Key)

# ✅ Loop through selected best models only
for (key in selected_keys) {

  model <- model_list[[key]]
  if (is.null(model)) next

  # --- Parse components ---
  parts <- strsplit(key, "_vs_")[[1]]
  response <- parts[1]
  predictor_and_type <- parts[2]
  predictor <- sub("_(GAM|GLM|TGAM)$", "", predictor_and_type)
  model_type <- sub("^.*_(GAM|GLM|TGAM)$", "\\1", predictor_and_type)

  # --- Build prediction grid ---
  x_min <- min(model$model[[predictor]], na.rm = TRUE)
  x_max <- max(model$model[[predictor]], na.rm = TRUE)

  if (!is.finite(x_min) || !is.finite(x_max) || x_min == x_max) next

  pred_vals <- seq(x_min, x_max, length.out = 300)
  pred_df <- setNames(data.frame(pred_vals), predictor)

  # --- TGAM hinge ---
  thr <- NA_real_
  if (model_type == "TGAM") {
    thr <- attr(model, "threshold")
    pred_df$hinge <- pmax(0, pred_df[[predictor]] - thr)
  }

  # --- Predictions ---
  pred_df$fit <- predict(model, newdata = pred_df, type = "response")

  pred_conf <- predict(model, newdata = pred_df, type = "link", se.fit = TRUE)
  pred_df$se <- pred_conf$se.fit

  pred_df$upper <- model$family$linkinv(pred_conf$fit + 1.96 * pred_df$se)
  pred_df$lower <- model$family$linkinv(pred_conf$fit - 1.96 * pred_df$se)

  # --- Colours ---
  glm_col  <- "blue"
  gam_col  <- "darkgreen"
  tgam_col <- "black"
  thr_col  <- "red"

  # --- Base plot ---
  p <- ggplot() +
    geom_point(
      data = model$model,
      aes(x = .data[[predictor]], y = .data[[response]]),
      alpha = 0.5
    ) +
    labs(
      # title = paste(response, "~", predictor),
      # subtitle = paste("Best model:", model_type),
      x = predictor,
      y = response
    ) +
    theme_minimal(base_size = 11)

  # =====================
  # GLM
  # =====================
  if (model_type == "GLM") {

    p <- p +
      geom_line(
        data = pred_df,
        aes(x = .data[[predictor]], y = .data[["fit"]]),
        color = glm_col, linewidth = 1
      ) +
      geom_ribbon(
        data = pred_df,
        aes(x = .data[[predictor]],
            ymin = .data[["lower"]],
            ymax = .data[["upper"]]),
        fill = glm_col, alpha = 0.2
      )
  }

  # =====================
  # GAM
  # =====================
  if (model_type == "GAM") {

    p <- p +
      geom_line(
        data = pred_df,
        aes(x = .data[[predictor]], y = .data[["fit"]]),
        color = gam_col, linewidth = 1
      ) +
      geom_ribbon(
        data = pred_df,
        aes(x = .data[[predictor]],
            ymin = .data[["lower"]],
            ymax = .data[["upper"]]),
        fill = gam_col, alpha = 0.2
      )

    # ✅ Derivative zero-crossings
    if (exists("crossings_df") && nrow(crossings_df) > 0) {
      crossings <- crossings_df$Derivative_Zero_Crossing[crossings_df$Model == key]

      if (length(crossings) > 0) {
        for (v in crossings) {
          if (is.finite(v)) {
            p <- p +
              geom_vline(xintercept = v, linetype = "dashed", color = thr_col)
          }
        }
      }
    }

    # ✅ Bootstrap slope shading
    if (exists("gam_slope_boot_df") && nrow(gam_slope_boot_df) > 0) {
      slope_dat <- gam_slope_boot_df %>% filter(Model == key)

      if (nrow(slope_dat) > 0) {
        xg <- slope_dat[[predictor]]
        sig <- with(slope_dat,
                    ifelse(slope_q025 > 0 | slope_q975 < 0, TRUE, FALSE))

        if (any(sig, na.rm = TRUE)) {
          p <- p +
            geom_ribbon(
              data = data.frame(x = xg, sig = sig),
              aes(x = x, ymin = -Inf, ymax = Inf),
              inherit.aes = FALSE,
              fill = "orange", alpha = 0.05
            )
        }
      }
    }
  }

  # =====================
  # TGAM
  # =====================
  if (model_type == "TGAM") {

    p <- p +
      geom_line(
        data = pred_df,
        aes(x = .data[[predictor]], y = .data[["fit"]]),
        color = tgam_col, linewidth = 1
      ) +
      geom_ribbon(
        data = pred_df,
        aes(x = .data[[predictor]],
            ymin = .data[["lower"]],
            ymax = .data[["upper"]]),
        fill = tgam_col, alpha = 0.2
      )

    # ✅ Threshold line
    if (!is.null(thr) && is.finite(thr)) {
      p <- p +
        geom_vline(xintercept = thr,
                   linetype = "dashed",
                   color = thr_col)
    }
  }

  plot_list[[key]] <- p
}

# ✅ Combine panels (ordered by pressure)

combined_plot <- wrap_plots(
  plot_list,
  ncol = length(unique(plot_df$Predictor)),
  byrow = FALSE   # ✅ THIS IS THE KEY LINE
)


# ✅ Save output
dir.create(here::here("figures"), showWarnings = FALSE, recursive = TRUE)

ggsave(
  filename = here::here("figures", paste0("bestmodel_", region, ".png")),
  plot = combined_plot,
  width = 20, height = 12,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

# split plots by indicator
```{r}

 # ✅ Define global indicator order (REQUIRED)
all_responses <- sort(unique(model_selection$Response))

# ✅ Split by pressure (rows)
plot_groups <- plot_df %>%
  group_split(Predictor)

grouped_plot_list <- list()

for (grp in plot_groups) {

  pred_name <- unique(grp$Predictor)

  # ✅ create full indicator grid
  subplots <- vector("list", length(all_responses))
  names(subplots) <- all_responses

  for (resp in all_responses) {

    row_match <- grp %>% filter(Response == resp)

    if (nrow(row_match) == 1) {

      key <- paste0(resp, "_vs_", pred_name, "_", row_match$Model_Type)

      if (key %in% names(plot_list)) {
        subplots[[resp]] <- plot_list[[key]]
      } else {
        subplots[[resp]] <- ggplot() + theme_void()
      }

    } else {
      # ✅ empty panel (kept)
      subplots[[resp]] <- ggplot() + theme_void()
    }
  }

  # ✅ one row per pressure
  p_group <- wrap_plots(
    subplots,
    ncol = length(all_responses),
    byrow = TRUE
  ) +
    plot_annotation(title = pred_name)

  grouped_plot_list[[pred_name]] <- p_group
}

# ✅ stack rows
combined_grouped_plot <- wrap_plots(
  grouped_plot_list,
  ncol = 1
)


ggsave(
  filename = here::here("figures", paste0("bestmodel_", region, ".png")),
  plot = combined_grouped_plot,
  width = 20,
  height = 2 * length(grouped_plot_list),  # ✅ dynamic height
  units = "in",
  dpi = 600,
  bg = "white"
)


```

# split plots by individual pressure
```{r}
# ---- Group plots by pressure and save ----

dir.create(
  here::here("figures", "by_pressure"),
  showWarnings = FALSE,
  recursive = TRUE
)

pressure_groups <- split(plot_df, plot_df$Predictor)

for (pressure_name in names(pressure_groups)) {

  grp <- pressure_groups[[pressure_name]]

  # Get corresponding plot keys
  keys <- grp$Model_Key

  # Extract plots
  pressure_plots <- plot_list[keys]

  # Remove NULLs just in case
  pressure_plots <- pressure_plots[!sapply(pressure_plots, is.null)]

  if (length(pressure_plots) == 0) next

  # Arrange plots for this pressure
  pressure_panel <- wrap_plots(
    pressure_plots,
    ncol = min(3, length(pressure_plots))
  ) +
    plot_annotation(
      title = pressure_name
    )

  # Dynamic height
  n_rows <- ceiling(length(pressure_plots) /
                      min(3, length(pressure_plots)))

  ggsave(
    filename = here::here(
      "figures",
      "by_pressure",
      paste0("pressure_", pressure_name, "_", region, ".png")
    ),
    plot = pressure_panel,
    width = 12,
    height = max(4, 4 * n_rows),
    units = "in",
    dpi = 600,
    bg = "white"
  )

  message("Saved: ", pressure_name)
}
```
 
# Step 9. Derivative panel grid (slope curve with CI)

```{r}
# ----  Optional derivative (slope) panels ----
deriv_plot_list <- list()
for (i in seq_len(nrow(model_selection))) {
  row <- model_selection[i, ]
  if (row$Model_Type != "GAM") next

  response   <- row$Response
  predictor  <- row$Predictor
  key        <- paste0(response, "_vs_", predictor, "_", "GAM")

  slope_dat <- gam_slope_boot_df %>% dplyr::filter(Model == key)
  if (nrow(slope_dat) == 0) next

  dp <- ggplot2::ggplot(slope_dat, aes_string(x = predictor, y = "slope_med")) +
    ggplot2::geom_ribbon(aes(ymin = slope_q025, ymax = slope_q975),
                         fill = "steelblue", alpha = 0.2) +
    ggplot2::geom_line(color = "steelblue4", linewidth = 0.9) +
    ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
    ggplot2::labs(title = paste0(response, " ~ ", predictor, " (slope)"),
                  x = predictor, y = "s′(x) [link scale]") +
    ggplot2::theme_minimal(base_size = 10)

  # Mark slope onset if available
  ssum <- gam_slope_boot_summary_df %>% dplyr::filter(Model == key)
  if (nrow(ssum) == 1 && is.finite(ssum$Slope_Onset_x)) {
    dp <- dp + ggplot2::geom_vline(xintercept = ssum$Slope_Onset_x, color = "orange", linetype = "dashed")
  }

  deriv_plot_list[[key]] <- dp
}

if (length(deriv_plot_list)) {
  deriv_combined_plot <- patchwork::wrap_plots(deriv_plot_list, ncol = 4)
  print(deriv_combined_plot)
  dir.create(here::here("figures"), showWarnings = FALSE, recursive = TRUE)
  ggplot2::ggsave(
    filename = here::here("figures", paste0("nafo_derivatives_", region, ".png")),
    plot     = deriv_combined_plot,
    width    = 16, height = 12, units = "in", dpi = 600, bg = "white"
  )
}
```

# Step 10. Save results
```{r}
# ---- Save results ----
dir.create(here::here("results"), showWarnings = FALSE, recursive = TRUE)

rdata_path <- here::here("results", paste0("ei_threshold_results_", region, ".RData"))

# ✅ Helper to safely create empty object if missing
ensure_object <- function(name, default) {
  if (!exists(name, envir = .GlobalEnv)) {
    assign(name, default, envir = .GlobalEnv)
  }
}

# ✅ Ensure ALL objects exist
ensure_object("threshold_df", tibble::tibble())
ensure_object("tgam_boot_summary_df", tibble::tibble())
ensure_object("tgam_boot_samples", list())
ensure_object("crossings_df", tibble::tibble())
ensure_object("effective_thresholds_df", tibble::tibble())
ensure_object("gam_effective_boot_summary_df", tibble::tibble())
ensure_object("gam_effective_boot_samples", list())
ensure_object("gam_slope_boot_df", tibble::tibble())
ensure_object("gam_slope_boot_summary_df", tibble::tibble())
ensure_object("unified_thresholds", tibble::tibble())
ensure_object("model_list", list())

# ✅ Now safe to save
save(
  best_models_table,
  model_selection,
  comparison_results,
  threshold_df,
  tgam_boot_summary_df,
  tgam_boot_samples,
  crossings_df,
  effective_thresholds_df,
  gam_effective_boot_summary_df,
  gam_effective_boot_samples,
  gam_slope_boot_df,
  gam_slope_boot_summary_df,
  unified_thresholds,
  model_list,
  file = rdata_path
)

message("Saved .RData: ", rdata_path)


```

#save as csv
```{r}


# ---- Save CSV versions of results ----

write.csv(
  model_selection,
  here::here("results", paste0("model_selection_", region, ".csv")),
  row.names = FALSE
)

write.csv(
  best_models_table,
  here::here("results", paste0("best_models_", region, ".csv")),
  row.names = FALSE
)

write.csv(
  unified_thresholds,
  here::here("results", paste0("unified_thresholds_", region, ".csv")),
  row.names = FALSE
)

write.csv(
  comparison_results,
  here::here("results", paste0("comparison_results_", region, ".csv")),
  row.names = FALSE
)

message("Saved CSV tables to: ", here::here("results"))


# ---- Save indicator and pressure datasets ----

write_csv(
  indicators,
  here::here("data_thresholds", paste0("indicators_", region, ".csv"))
)

write_csv(
  pressures,
  here::here("data_thresholds", paste0("pressures_", region, ".csv"))
)

message("Saved indicators and pressures CSV files.")


```

